### 2017-01-23 paper.tex
whitepaper outline

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

%For final version of Special issue article: For each paper in the special issue, up to 20 single column double-spaced pages, 11 point font size, including figures, tables and references. The total number of figures and tables may be up to 10 (sub-figures in (a), (b), (c), etc. counted separately). The total number of references may be up to 30. There should be at least 1.25" margin on left and right sides, and 1" margin from top and bottom. The figures and tables should be placed in the center of the column, and not tightly embedded into the text column. For some special issues, the page limit may be even less if more papers are included. In any event, check with the guest editors before writing the full manuscript.

% White paper due Feb 20, up to 4 pages length
% more info: https://signalprocessingsociety.org/sites/default/files/uploads/special_issues_deadlines/SPM_SI_geo_sp.pdf

\title{*Catchy short title*: \\
automated noise processing for near-surface characterization with seismic data from a fiber network}
\author{Eileen R. Martin, Fantine Huot/Yinbin Ma(*you guys figure out order*), Biondo L. Biondi, Steve Cole, Martin Karrenbach}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle

\begin{abstract} 
*****abstract words here *****
\end{abstract}

\section*{Motivation}
****what's ambient noise used for?****

****why use DAS? brief overview of its use****

\section*{New Contributions}
****what's new that we're doing, how it's different from what's been done before****

\section*{Background Theory}
****ambient noise background****

\section*{Experimental Overview}
****overview of the array, potential noise sources, ambient noise field****

\section*{Data Processing}
***overview of data processing workflow****

\subsection*{Estimating Sensor Coupling Response}
***talk about different responses from different sensors and how to compensate for these, use many earthquakes for relative responses****

\subsection*{Localizing Persistent Noise Sources}
***talk about localizing these things (motors, traffic, etc....) and show a little beamforming in small 1D parts of the array bandpassed to these persistent frequency peaks***

\subsection*{Removing Earthquakes and Quarry Blasts}
****talk about getting rid of big events lighting up whole array****

\subsection*{Removing Vehicle Noise}
****talk about clustering and comparing noise sources-including rotations and stretch of example noise****

\subsection*{Removing Correlated Noise Sources}
*****talk about detecting temporally and spatially correlated noise sources, more generally than cars****

\subsection*{Traveltime Picking}
*** talk about traveltime picking before Rayleigh wave tomography****

\subsection*{Tomography}
**** talk about how near-surface tomography is set up****

\section*{Results}
***overview of results****

\subsection*{Convergence Rate Improvement Due to Pre-Processing}
***show comparisons of cross-correlations with/without sensor coupling response, and with scaling by energy per channel of cross-correlations (also show spectrum per channel of these)****
\\
******show comparisons of convergence rates with/without vehicles removed, and with vehicles just removed via STA/LTA jumps****

\subsection*{Tomography Along Array Lines}
***show tomography results along multiple lines on campus****

\section*{Summary}
****summary words****

\begin{thebibliography}{10}
\bibitem{somebibitem} ***first bibitem***

\end{thebibliography}

\end{document}

### 2017-02-14 paper.tex
Word count, paragraph assignment, figure count

@@
 \begin{abstract} 
-*****abstract words here *****
+*****abstract words here - 150-200 Eileen*****
 \end{abstract}
@@
 \section*{Motivation}
-****what's ambient noise used for?****
+****what's ambient noise used for? - 200-250 Eileen****
 
@@
 \section*{New Contributions}
-****what's new that we're doing, how it's different from what's been done before****
+****what's new that we're doing, how it's different from what's been done before - 150 Eileen, little Fantine & Yinbin****
 
 \section*{Background Theory}
-****ambient noise background****
+****ambient noise background - 150-220 Eileen****
 
 \section*{Experimental Overview}
-****overview of the array, potential noise sources, ambient noise field****
+****overview of the array, potential noise sources, ambient noise field - 200 Eileen****
+****Figure : 2 panels: Installation + map****
 
 \section*{Data Processing}
-***overview of data processing workflow****
+***overview of data processing workflow - 200-250 Eileen****
+***Figure : stft plots at a few different spots***
+***Mention whitening here***
+***Mention earthqukes and quarry blasts here***
 
@@
 ***talk about different responses from different sensors and how to compensate for these, use many earthquakes for relative responses****
+***100 Eileen***
+***Figure : Boxed earthquake recording + response versus channel***
 
-\subsection*{Localizing Persistent Noise Sources}
-***talk about localizing these things (motors, traffic, etc....) and show a little beamforming in small 1D parts of the array bandpassed to these persistent frequency peaks***
+% \subsection*{Localizing Persistent Noise Sources}
+% ***talk about localizing these things (motors, traffic, etc....) and show a little beamforming in small 1D parts of the array bandpassed to these persistent frequency peaks***
 
-\subsection*{Removing Earthquakes and Quarry Blasts}
-****talk about getting rid of big events lighting up whole array****
+% \subsection*{Removing Earthquakes and Quarry Blasts}
+% ****talk about getting rid of big events lighting up whole array****
 
 \subsection*{Removing Vehicle Noise}
-****talk about clustering and comparing noise sources-including rotations and stretch of example noise****
+****talk about clustering and comparing noise sources****
+***500 Fantine & Yinbin***
+***Figures : Before/after car mute + clusters and projection on map***
+***Little Eileen for sentence about why cars are a problem***
 
-\subsection*{Removing Correlated Noise Sources}
-*****talk about detecting temporally and spatially correlated noise sources, more generally than cars****
+% \subsection*{Removing Correlated Noise Sources}
+% *****talk about detecting temporally and spatially correlated noise sources, more generally than cars****
 
-\subsection*{Traveltime Picking}
-*** talk about traveltime picking before Rayleigh wave tomography****
+% \subsection*{Traveltime Picking}
+% *** talk about traveltime picking before Rayleigh wave tomography****
+% *** STANDBY ***
 
-\subsection*{Tomography}
-**** talk about how near-surface tomography is set up****
+% \subsection*{Tomography}
+% **** talk about how near-surface tomography is set up****
+% *** STANDBY ***
 
@@
 ***overview of results****
+*** Convergence rates
+***150 Eileen *** 
+***Figure : line plot of correlation coeeficients vs epoch/time***
 
-\subsection*{Convergence Rate Improvement Due to Pre-Processing}
+% \subsection*{Convergence Rate Improvement Due to Pre-Processing}
 ***show comparisons of cross-correlations with/without sensor coupling response, and with scaling by energy per channel of cross-correlations (also show spectrum per channel of these)****
@@
 
-\subsection*{Tomography Along Array Lines}
-***show tomography results along multiple lines on campus****
+% \subsection*{Tomography Along Array Lines}
+% ***show tomography results along multiple lines on campus****
+% *** STANDBY ***
 
@@
 ****summary words****
+***100 Eileen ***
 

### 2017-02-14 paper.tex
formatting change to white paper

--- 2017-01-23 paper.tex
@@
 \documentclass[11pt]{amsart}
-\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
+\usepackage[margin=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
 \geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
@@
 
-\title{*Catchy short title*: \\
-automated noise processing for near-surface characterization with seismic data from a fiber network}
-\author{Eileen R. Martin, Fantine Huot/Yinbin Ma(*you guys figure out order*), Biondo L. Biondi, Steve Cole, Martin Karrenbach}
+\title{****title*****}
+\author{Eileen R. Martin, Fantine Huot, Yinbin Ma, Robert Cieplicki, Biondo L. Biondi}
 %\date{}                                           % Activate to display a given date or no date
@@
 
-\begin{abstract} 
-*****abstract words here *****
-\end{abstract}
+\section*{Motivation}  *****2/3 page Eileen****
 
-\section*{Motivation}
-****what's ambient noise used for?****
+\section*{Significance} ******2/3 page Eileen****
 
-****why use DAS? brief overview of its use****
+\section*{Outline of the paper}
+	\begin{enumerate}
+	\item Abstract/summary
+	\item Introduction 
+		\begin{itemize}
+		\item *****words about intro, motivation****
+		\end{itemize}
+	\item Related work
+		\begin{enumerate}
+		\item Background on ambient noise: how it works that for the cost of just laying down receivers, you can extract signals mimicking those of an active seismic survey with a source at any receiver ****add citations******
+		\item Background on Distributed Acoustic Sensing: how it works ****add citations****, how it has been used so far ****add citations****
+		\item Examples of ambient seismic noise studies using Distributed Acoustic Sensing *****add citations****
+		\end{enumerate}
+	\item Data Processing: we process the noise we record so that it mimicks data that would be recorded if the noise field followed theoretical ideals.
+Some common strategies include ****citations to review like papers*****
+		\begin{enumerate}
+		\item Clustering of data to speed up noise field exploration
+		\item How to use earthquakes to estimate sensor coupling 
+		\item Detection of localized, persistent noise sources 
+		\item Removing vehicle noise: design of filter for cars, for each piece of noise find nearest cluster to see if it's a car, then filter out the car while preserving the use of other noise in the same time and place
+		\end{enumerate}
+	\item Results
+		\begin{enumerate}
+		\item Comparison of convergence rates of virtual source response function estimates ****cite Seats 2012**** (both cross-correlation and cross-coherence) for traditional processing and new processing methods
+		\item Velocity model resulting from ambient noise tomography
+		\end{enumerate}
+	\item Conclusions
+	\item Acknowledgments
+	\end{enumerate}
 
-\section*{New Contributions}
-****what's new that we're doing, how it's different from what's been done before****
+%\begin{abstract} 
+%*****abstract words here - 150-200 Eileen*****
+%\end{abstract}
 
-\section*{Background Theory}
-****ambient noise background****
+%\section*{Motivation}
+%****what's ambient noise used for? - 200-250 Eileen****
 
-\section*{Experimental Overview}
-****overview of the array, potential noise sources, ambient noise field****
+%****why use DAS? brief overview of its use****
 
-\section*{Data Processing}
-***overview of data processing workflow****
+%\section*{New Contributions}
+%****what's new that we're doing, how it's different from what's been done before - 150 Eileen, little Fantine \& Yinbin****
 
-\subsection*{Estimating Sensor Coupling Response}
-***talk about different responses from different sensors and how to compensate for these, use many earthquakes for relative responses****
+%\section*{Background Theory}
+%****ambient noise background - 150-220 Eileen****
 
-\subsection*{Localizing Persistent Noise Sources}
-***talk about localizing these things (motors, traffic, etc....) and show a little beamforming in small 1D parts of the array bandpassed to these persistent frequency peaks***
+%\section*{Experimental Overview}
+%****overview of the array, potential noise sources, ambient noise field - 200 Eileen****
+%****Figure : 2 panels: Installation + map****
 
-\subsection*{Removing Earthquakes and Quarry Blasts}
-****talk about getting rid of big events lighting up whole array****
+%\section*{Data Processing}
+%***overview of data processing workflow - 200-250 Eileen****
+%***Figure : stft plots at a few different spots***
+%***Mention whitening here***
+%***Mention earthqukes and quarry blasts here***
 
-\subsection*{Removing Vehicle Noise}
-****talk about clustering and comparing noise sources-including rotations and stretch of example noise****
+%\subsection*{Estimating Sensor Coupling Response}
+%***talk about different responses from different sensors and how to compensate for these, use many earthquakes for relative responses****
+%***100 Eileen***
+%***Figure : Boxed earthquake recording + response versus channel***
 
-\subsection*{Removing Correlated Noise Sources}
-*****talk about detecting temporally and spatially correlated noise sources, more generally than cars****
+%% \subsection*{Localizing Persistent Noise Sources}
+%% ***talk about localizing these things (motors, traffic, etc....) and show a little beamforming in small 1D parts of the array bandpassed to these persistent frequency peaks***
 
-\subsection*{Traveltime Picking}
-*** talk about traveltime picking before Rayleigh wave tomography****
+%% \subsection*{Removing Earthquakes and Quarry Blasts}
+%% ****talk about getting rid of big events lighting up whole array****
 
-\subsection*{Tomography}
-**** talk about how near-surface tomography is set up****
+%\subsection*{Removing Vehicle Noise}
+%****talk about clustering and comparing noise sources****
+%***500 Fantine \& Yinbin***
+%***Figures : Before/after car mute + clusters and projection on map***
+%***Little Eileen for sentence about why cars are a problem***
 
-\section*{Results}
-***overview of results****
+%% \subsection*{Removing Correlated Noise Sources}
+%% *****talk about detecting temporally and spatially correlated noise sources, more generally than cars****
 
-\subsection*{Convergence Rate Improvement Due to Pre-Processing}
-***show comparisons of cross-correlations with/without sensor coupling response, and with scaling by energy per channel of cross-correlations (also show spectrum per channel of these)****
+%% \subsection*{Traveltime Picking}
+%% *** talk about traveltime picking before Rayleigh wave tomography****
+%% *** STANDBY ***
+
+%% \subsection*{Tomography}
+%% **** talk about how near-surface tomography is set up****
+%% *** STANDBY ***
+
+%\section*{Results}
+%***overview of results****
+%*** Convergence rates
+%***150 Eileen *** 
+%***Figure : line plot of correlation coeeficients vs epoch/time***
+
+%% \subsection*{Convergence Rate Improvement Due to Pre-Processing}
+%***show comparisons of cross-correlations with/without sensor coupling response, and with scaling by energy per channel of cross-correlations (also show spectrum per channel of these)****
+%\\
+%******show comparisons of convergence rates with/without vehicles removed, and with vehicles just removed via STA/LTA jumps****
+
+%% \subsection*{Tomography Along Array Lines}
+%% ***show tomography results along multiple lines on campus****
+%% *** STANDBY ***
+
+%\section*{Summary}
+%****summary words****
+%***100 Eileen ***
+
+\section*{Biographies}
+\textbf{Eileen R.
+Martin} \texttt{ermartin@stanford.edu} is a Ph.D.
+student in Stanford's Institute for Computational and Mathematical Engineering, and an M.S.
+student in the geophysics department.
+She first started working with ambient noise and DAS data from permafrost thaw monitoring experiments done by Lawrence Berkeley National Lab and the US Army Corps of Engineers.
 \\
-******show comparisons of convergence rates with/without vehicles removed, and with vehicles just removed via STA/LTA jumps****
-
-\subsection*{Tomography Along Array Lines}
-***show tomography results along multiple lines on campus****
-
-\section*{Summary}
-****summary words****
+\textbf{Fantine Huot} \texttt{**email address} *****
+\\
+\textbf{Yinbin Ma} \texttt{**email address} *****
+\\
+\textbf{Robert Cieplicki} \texttt{***email address} ****
+\\
+\textbf{Biondo L. Biondi} \texttt{**email address} ****
 

### 2017-02-15 paper.tex
adding references, still have space for 10 more

@@
-\documentclass[11pt]{amsart}
+\documentclass[11pt]{article}
 \usepackage[margin=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
@@
 \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
+\usepackage{authblk}
 
@@
 \title{****title*****}
-\author{Eileen R. Martin, Fantine Huot, Yinbin Ma, Robert Cieplicki, Biondo L. Biondi}
+\author[1,2]{Eileen R. Martin}
+\author[2]{Fantine Huot}
+\author[1]{Yinbin Ma}
+\author[2]{Robert Cieplicki}
+\author[1,2]{Biondo L. Biondi}
+\affil[1]{Institute for Computational and Mathematical Engineering, Stanford University}
+\affil[2]{Department of Geophysics, Stanford University}
 %\date{}                                           % Activate to display a given date or no date
@@
 \\
+\textbf{Steve Cole} \texttt{Steve.Cole@optasense.com} is Manager of Integrated Analysis at OptaSense.
+He has worked in the oil and gas industry for over 30 years at companies including Fugro, Chevron, and 4th Wave Imaging, which he co-founded.
+He has a Ph.D.
+in geophysics from Stanford.
+\\
 \textbf{Biondo L. Biondi} \texttt{**email address} ****
 
-\begin{thebibliography}{10}
-\bibitem{somebibitem} ***first bibitem***
+\begin{thebibliography}{30} % no more than 30 references
+
+% example of trenched ambient noise recording with distributed acoustic sensing
+\bibitem{AjoFranklin2015} J.
+Ajo-Franklin, N.
+Lindsey, S.
+Dou, T.M.
+Daley, B.
+Freifeld, E.R.
+Martin, M.
+Robertson, C.
+Ulrich and A.
+Wagner, "A Field Test of Distributed Acoustic Sensing for Ambient Noise Recording," \textit{SEG Technical Program Expanded Abstracts}, pp.
+2620-2624, 2015.
+
+% working with DAS data, how DAS works basically
+\bibitem{Bakku2015} S.K. Bakku, "Fracture Characterization from Seismic Measurements in a Borehole," Ph.D. Thesis, Massachusetts Institute of Technology, 2015.
+
+% temporal and spectral and one bit normalization
+\bibitem{Bensen2007} G.D.
+Bensen, M.H.
+Ritzwoller, M.P.
+Barmin, A.L.
+Levshin, F.
+Lin, M.P.
+Moschetti, N.M.
+Shapiro and Y.
+Yang, "Processing seismic ambient noise data to obtain reliable broad-band surface wave dispersion measurements," \textit{Geophys.
+J.
+Int.}, vol.
+169, no.
+****, pp.
+1239-1260, 2007.
+
+% whitenging for robustness to velocity changes
+\bibitem{Daskalakis2016} E.
+Daskalakis, C.P.
+Evangelidis, J.
+Garnier, N.S.
+Melis, G.
+Papanicolaou and C.
+Tsoga, "Robust seismic velocity change estimation using ambient noise recordings," \textit{Geophys.
+J.
+Int.}, vol.
+205, no.
+3, pp.
+1926-1936, 2016.
+
+% ambient noise with mine blasts (similar to earthquakes)
+\bibitem{Girard2016} A.J. Girard and J. Shragge, "Extracting body waves from ambient seismic recordings," \textit{SEG Technical Program Expanded Abstracts}, pp. 2715-2719, 2016.
+
+% sensitivity of DAS cables with slippage, straight or helical
+\bibitem{Kuvshinov2015} B.N. Kuvshinov, "Interaction of helically wound fibre-optic cables with plane seismic waves," \textit{Geophysical Prospecting}, vol. ***, no. ***, pp. 1-18, 2015.
+
+% using horizontal components in interferometry as Rayleigh and Love wave components
+\bibitem{Lin2008} F.
+Lin, M.P.
+Moschetti and M.H.
+Ritzwoller, "Surface wave tomography of the western United States from ambient seismic noise: Rayleigh and Love wave phase velocity maps," \textit{Geophys.
+J.
+Int.} vol.
+173, no.
+***, pp.
+281-298, 2008.
+
+% example of interferometry of ambient noise recorded by a trenched DAS array
+\bibitem{Martin2015} E.R.
+Martin, J.
+Ajo-Franklin, S.
+Dou, N.
+Lindsey, T.M.
+Daley, B.
+Freifeld, M.
+Robertson, A.
+Wagner, C.
+Ulrich, "Interferometry of ambient noise from a trenched distributed acoustic sensing array," \textit{SEG Technical Program Expanded Abstracts}, pp.
+2445-2450, 2015.
+
+% example of roadside DAS interferometry
+\bibitem{Martin2016} E.R.
+Martin, N.J.
+Lindsey, S.
+Dou, J.
+Ajo-Franklin, T.M.
+Daley, B.
+Freifeld, M.
+Robertson, C.
+Ulrich, A.
+Wagner and K.
+Bjella, "Interferometry of a roadside DAS array in Fairbanks, AK," \textit{SEG Technical Program Expanded Abstracts}, pp.
+2725-2629, 2016.
+
+% cross correlation, coherence and deconvolution, and using traffic noise
+\bibitem{Nakata2011} N.
+Nakata, R.
+Snieder, T.
+Tsuji, K.
+Larner and T.
+Matsuoka, "Shear wave imaging from traffic noise using seismic interferometry by cross-coherence," \textit{Geophysics}, vol.
+76, no.
+6, pp.
+SA97-SA106, 2011.
+
+% ambient noise on dense array in Long Beach
+\bibitem{Nakata2015} N.
+Nakata, J.P.
+Chang, J.F.
+Lawrence and P.
+Bou\'{e}, "Body wave extraction and tomography at Long Beach, California, with ambient-noise interferometry," \textit{J.
+Geophys.
+Res.
+Solid Earth}, vol.
+120, pp.
+1159-1173, 2015.
+
+% how DAS works
+\bibitem{Posey2000} R.
+Posey Jr., G.A.
+Johnson and S.T.
+Vohra, "Strain sensing based on coherence Rayleigh scattering in an optical fibre," \textit{Electronics Letters}, vol.
+36, no.
+20, pp.
+1688-1689, 2000.
+
+% overlapping windows, method of measuring convergence with correlation coefficient
+\bibitem{Seats2012} K.J. Seats, J.F. Lawrence and G.A. Prieto, "Improved ambient noise correlation functions using Welch's method," \textit{Geophys. J. Int.}, vol. 188, no. ***, pp. 513-523, 2012.
+
+% fundamental paper in ambient noise
+\bibitem{Lobkis2001} O.I. Lobkis and R.L. Weaver, "On the emergence of the Green's function in the correlations of a diffuse field," \textit{J. Acoust. Soc. Am.}, vol. 110, no. 6, pp. 3011-3017, 2001.
+
+% DAS conversion to velocities, and VSP example and side-by-side multi-mode, single-mode and 3C geophones
+\bibitem{Miller2016} D.E.
+Miller, T.M.
+Daley, B.M.
+Freifeld, M.
+Robertson, J.
+Cocker and M.
+Craven, "Simultaneous Acquisition of Distributed Acoustic Sensing VSP with Multi-mode and Single-mode FIber-optic Cables and 3C-Geophones at the Aquistore CO$_2$ Storage Site," \textit{CSEG Recorder}, pp.
+28-***, June 2016.
+
+% basics of interferometry
+\bibitem{Wapenaar2010A} K.
+Wapenaar, D.
+Draganov, R.
+Snieder, X.
+Campman and A.
+Verdel, "Tutorial on seismic interferometry: Part 1 - Basic principles and applications," \textit{Geophysics}, vol.
+75, no.
+5, pp.
+75A195-75A209, 2010.
+% more basics of interferometry
+\bibitem{Wapenaar2010B} K.
+Wapenaar, E.
+Slob, R.
+Snieder and A.
+Curtis, "Tutorial on seismic interferometry: Part 2 - Underlying theory and new advances," \textit{Geophysics}, vol.
+75, no.
+5, pp.
+75A211-75A227, 2010.
+
+% example of DAS for microseismic event detection
+\bibitem{Webster2013} P.
+Webster, J.
+Wall, C.
+Perkins and M.
+Molenaar, "Micro-Seismic Detection using Distributed Acoustic Sensing," \textit{SEG Technical Program Expanded Abstracts}, pp.
+2459-2463, 2013.
+
+% another example of DAS interferometry, this one side-by-side with geophones
+\bibitem{Zeng2017} X.
+Zeng, C.
+Thurber, H.
+Wang, D.
+Fratta, E.
+Matzel, PoroTomo Team, "High-resolution Shallow Structure Revealed with Ambient Noise Tomography on a Dense Array," \textit{Proceedings, 42nd Workshop on Geothermal Reservoir Engineering}, Stanford University, Stanford, CA, Feb.
+13-15, 2017, SGP-TR-212.
+
+% velocity artifacts from ambient noise source changes
+\bibitem{Zhan2013} Z.
+Zhan, V.C.
+Tsai and R.W.
+Clayton, "Spurious velocity changes caused by temporal variations in ambient noise frequency content," \textit{Geophys.
+J.
+Int}, vol.
+194, no.
+***, pp.
+1574-1581, 2013.
 

### 2017-02-16 paper.tex
did a little more motivation and significance

@@
 \usepackage{authblk}
+\usepackage{titlesec}
 
@@
 
-\title{****title*****}
-\author[1,2]{Eileen R. Martin}
+% KEYWORDS from IEEE Taxonomy 2017:
+% Sensors: Acoustic Sensors
+% Sensors: Optical fiber sensors
+% Signal processing: Acoustic signal processing
+% Signal processing: Array signal processing
+% Signal processing: Geophysical signal processing
+% Signal processing: Multidimensional signal processing
+% Science-general: Waves, Seismic waves
+% Science-general: Geophysics, Surface waves
+
+\title{\vspace{-1.8cm} \Large Demystifying Ambient Seismic Noise Processing: \\Automating source analysis and processing decisions in the presence of coherent anthropogenic, natural, and sensor noise \vspace{-0.4cm}}
+\author[1,2]{\small Eileen R. Martin}
 \author[2]{Fantine Huot}
@@
 \author[2]{Robert Cieplicki}
-\author[1,2]{Biondo L. Biondi}
-\affil[1]{Institute for Computational and Mathematical Engineering, Stanford University}
-\affil[2]{Department of Geophysics, Stanford University}
-%\date{}                                           % Activate to display a given date or no date
+\author[3]{Steve Cole}
+\author[3]{Martin Karrenbach}
+\author[1,2]{Biondo L. Biondi \vspace{-0.3cm}}
+\affil[1]{Institute for Computational and Mathematical Engineering, Stanford University, Stanford, CA}
+\affil[2]{Department of Geophysics, Stanford University, Stanford, CA}
+\affil[3]{OptaSense, Brea, CA}
+\date{}
 
@@
 
-\section*{Motivation}  *****2/3 page Eileen****
+\vspace{-1.8cm}
+\subsection*{Motivation} 
+\vspace{-0.2cm}
+****motivation for near surface seismic imaging****
+\\
+\\
+****motivation for using DAS****history \& current issues in DAS****
+Distributed acoustic sensing (DAS) probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors.
+It does this by performing optical interferometry on Rayleigh backscattered photons, associating the resulting phase shift with a subset of cable based on the two-way travel time of those photons, then applying a linear transformation to the change in phase shift associated with that subset of cable \cite{Posey2000}, \cite{Daley2013}, \cite{Bakku2015}.
+That cable can be either single- or multi-mode \cite{Miller2016}.
+For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber inside the cable to make its angular sensitivity more ideal for a variety of source-receiver geometries \cite{Kuvshinov2016}.
+\\
+\\
+***motivation for using ambient noise****current issues in ambient noise*****
+\\
+\\
+There have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}.
+This type of installation allows ****
+By running fibers in existing telecommunications conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor are becoming a reality.
+When the interrogator unit is probing a section of fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes.
 
-\section*{Significance} ******2/3 page Eileen****
+\vspace{-0.3cm}
+\subsection*{Significance} 
+\vspace{-0.2cm}
+This manuscript will walk through the basics (and some new tools) for processing ambient noise to do near-surface seismic imaging. Its purposes:
+\vspace{-0.2cm}
+\begin{itemize}
+\item overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with ambient noise in the presence of non-ideal noises
+\vspace{-0.2cm}
+\item show optical engineers an application where an optical measurement device's noise level and sensitivity should be improved
+\vspace{-0.2cm}
+\item expose gepohysicists to an emerging acquisition method and a few machine learning tools and metrics to quantify processing decisions, a task that will become increasingly more difficult as data acquisition continues to become easier and cheaper
+\end{itemize}
+\vspace{-0.2cm}
+It is difficult to gain intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics loosely laying in existing telecommunications conduits undernear the Stanford University campus.
+This particular case study is an ideal sandbox for a tutorial that demonstrates the wide variety of issues on the near horizon as we push for broader application of ambient noise studies.
+\\
+\\
+We will discuss the noise level and sensitivities particular to recording with DAS, and a novel deployment that can greatly reduce costs (increasing data volumes): fiber loosely laying in existing telecommunications conduits.
+The array detects a wide variety of seismic noise sources that do not conform to the ideals of existing ambient noise theory: it sits in a seismically active region, 20 km from the Pacific ocean, 7 km from the San Francisco bay, with major highways on either side, a variety of roads with differing levels of traffic near the fiber, regular quarry blasts within 15 km, plumbing and HVAC systems throughout the site, multiple construction sites near the array, and foot and bicycle traffic throughout.
+With over 600 sensors continuously recording 50 samples per second since September 2016, manual inspection of most data is infeasible, making the automation tools shown in this paper critical to extracting subsurface information from these data.
 
-\section*{Outline of the paper}
+\vspace{-0.3cm}
+\subsection*{Outline of the paper}
 	\begin{enumerate}
+	\vspace{-0.2cm}
 	\item Abstract/summary
-	\item Introduction 
-		\begin{itemize}
-		\item *****words about intro, motivation****
-		\end{itemize}
+	\vspace{-0.2cm}
+	\item Introduction: why we care about near surface imaging, discuss tradeoffs of sensor quality with cost and array density (DAS or MEMS, etc...), introduce idea of ambient noise to get rid of sources, ***
+	\vspace{-0.2cm}
 	\item Related work
 		\begin{enumerate}
+		\vspace{-0.2cm}
 		\item Background on ambient noise: how it works that for the cost of just laying down receivers, you can extract signals mimicking those of an active seismic survey with a source at any receiver ****add citations******
+		\vspace{-0.2cm}
 		\item Background on Distributed Acoustic Sensing: how it works ****add citations****, how it has been used so far ****add citations****
+		\vspace{-0.2cm}
 		\item Examples of ambient seismic noise studies using Distributed Acoustic Sensing *****add citations****
 		\end{enumerate}
+	\vspace{-0.2cm}
 	\item Data Processing: we process the noise we record so that it mimicks data that would be recorded if the noise field followed theoretical ideals.
@@
 		\begin{enumerate}
-		\item Clustering of data to speed up noise field exploration
-		\item How to use earthquakes to estimate sensor coupling 
-		\item Detection of localized, persistent noise sources 
-		\item Removing vehicle noise: design of filter for cars, for each piece of noise find nearest cluster to see if it's a car, then filter out the car while preserving the use of other noise in the same time and place
+		\vspace{-0.2cm}
+		\item Clustering/separation of data (k-means, hierarchical agglomerative, support vector machine separation) to speed up noise field exploration, identification of potential issues
+		\vspace{-0.2cm}
+		\item How to use earthquakes to estimate sensor coupling heterogeneity when array installation is relatively uncontrolled (specifically, a fiber loosely laying in a variety of conduits) 
+		\vspace{-0.2cm}
+		\item Automatically detect \& filter persistent noise sources from short-time Fourier transforms
+		\vspace{-0.2cm}
+		\item Automating the design of f-k domain filters for particular clusters based on separation from other clusters or background level noise.
+In particular, a demonstration of removing vehicle noise: time domain mute vs.
+manually designed f-k vs.
+automated design.
 		\end{enumerate}
-	\item Results
-		\begin{enumerate}
-		\item Comparison of convergence rates of virtual source response function estimates ****cite Seats 2012**** (both cross-correlation and cross-coherence) for traditional processing and new processing methods
-		\item Velocity model resulting from ambient noise tomography
-		\end{enumerate}
-	\item Conclusions
-	\item Acknowledgments
+	\vspace{-0.2cm}
+	\item Measuring improvements: measure convergence rates of virtual source response function estimates like \cite{Seats2012}, but  (both cross-correlation and cross-coherence) for traditional processing and new processing methods
+	\vspace{-0.2cm}
+	\item Show a few dispersion images and a velocity model resulting from surface wave tomography, a couple of the possible uses of the virtual source response estimates
+	\vspace{-0.2cm}
+	\item Conclusions and future directions
 	\end{enumerate}
@@
 %***100 Eileen ***
-
-\section*{Biographies}
+\vspace{-0.3cm}
+\subsection*{Biographies}
+\vspace{-0.2cm}
 \textbf{Eileen R.
-Martin} \texttt{ermartin@stanford.edu} is a Ph.D.
+Martin} (\texttt{ermartin@stanford.edu}) is a Ph.D.
 student in Stanford's Institute for Computational and Mathematical Engineering, and an M.S.
-student in the geophysics department.
+student in the Geophysics Department.
 She first started working with ambient noise and DAS data from permafrost thaw monitoring experiments done by Lawrence Berkeley National Lab and the US Army Corps of Engineers.
 \\
-\textbf{Fantine Huot} \texttt{**email address} *****
+\textbf{Fantine Huot} \texttt{fantine@sep.stanford.edu} received her M.S.
+in Science and Engineering from the Ecole Normale Sup\'erieure des Mines de Paris in 2013.
+Following graduation, she worked as a software developer in France.
+In fall 2015, she started her Ph.D.
+in Geophysics at Stanford University and joined the Stanford Exploration Project.
+She works on addressing earth science problems by leveraging machine learning algorithms.
 \\
@@
 \\
-\textbf{Robert Cieplicki} \texttt{***email address} ****
+\textbf{Robert Cieplicki} (\texttt{robcie@stanford.edu}) is a PhD student in the Zoback Stress and Crustal Mechanics Research Group (Geophysics Department, Stanford University).
+He applies machine learning to horizontal wells drilled in shale hydrocarbon reservoirs to predict the efficiency of hydraulic fracturing.
 \\
-\textbf{Steve Cole} \texttt{Steve.Cole@optasense.com} is Manager of Integrated Analysis at OptaSense.
+\textbf{Steve Cole} (\texttt{Steve.Cole@optasense.com}) is Manager of Integrated Analysis at OptaSense.
 He has worked in the oil and gas industry for over 30 years at companies including Fugro, Chevron, and 4th Wave Imaging, which he co-founded.
@@
 \\
+\textbf{Martin Karrenbach} (\texttt{***email address}) *******
+\\
 \textbf{Biondo L. Biondi} \texttt{**email address} ****
 
+\vspace{-0.3cm}
+\begingroup
+\titleformat*{\section}{\large \bf}
 \begin{thebibliography}{30} % no more than 30 references
-
+\vspace{-0.2cm}
+\small
 % example of trenched ambient noise recording with distributed acoustic sensing
@@
 2620-2624, 2015.
-
+\vspace{-0.2cm}
 % working with DAS data, how DAS works basically
 \bibitem{Bakku2015} S.K. Bakku, "Fracture Characterization from Seismic Measurements in a Borehole," Ph.D. Thesis, Massachusetts Institute of Technology, 2015.
-
+\vspace{-0.2cm}
 % temporal and spectral and one bit normalization
@@
 1239-1260, 2007.
-
+\vspace{-0.2cm}
+% overview of the array, reasons to use DAS in existing telecomm conduits
+\bibitem{Martin2017} E.R.
+Martin, B.L.
+Biondi, M.
+Karrenbach, S.
+Cole, "Continuous Subsurface Monitoring by Passice Seismic with Distributed Acoustic Sensors- The "Stanford Array" Experiment," \textit{Proceedings of the First EAGE Workshop on Practical Reservoir Monitoring}, 2017.
+\vspace{-0.2cm}
 % whitenging for robustness to velocity changes
@@
 1926-1936, 2016.
-
+\vspace{-0.2cm}
+% nice intro to DAS with a seismic active experiment
+\bibitem{Daley2013} T.M.
+Daley, B.M.
+Freifeld, J.
+Ajo-Franklin and S.
+Dou, "Field testing of fiber-optic distributed acoustic sensing (DAS) for subsurface seismic monitoring," \textit{The Leading Edge}, pp.
+936-942, June 2013.
+\vspace{-0.2cm}
+% an overview comparison of MEMS accelerometers
+\bibitem{Evans2014} J.R.
+Evans, R.M.
+Allen, A.I.
+Chung, E.S.
+Cochran, R.
+Guy, M.
+Hellweg and J.F.
+Lawrence, "Performance of Several Low-Cost Accelerometers," \textit{Seismol.
+Res.
+Lett.}, vol.
+85, no.
+1, pp.
+148-158, 2014.
+\vspace{-0.2cm}
+% another example of passive seismic data being analyzed with machine learning techniques (including CWT as important feature)
+\bibitem{Fisher2016} W.D.
+Fisher, T.K.
+Camp and V.V.
+Krzhizhanovskaya, "Anomaly detection in earth dam and levee passive seismic data using support vector machines and automatic feature selection," \textit{J.
+Comput.
+Sci.}, in press, doi: 10.1016/j.jocs.2016.11.016, 2016.
+\vspace{-0.2cm}
 % ambient noise with mine blasts (similar to earthquakes)
 \bibitem{Girard2016} A.J. Girard and J. Shragge, "Extracting body waves from ambient seismic recordings," \textit{SEG Technical Program Expanded Abstracts}, pp. 2715-2719, 2016.
-
+\vspace{-0.2cm}
+% lots of basic data analysis tools
+\bibitem{Hastie2009} T. Hastie, R. Tibshirani and J. Friedman, \textit{The Elements of Statistical Learning: Data Mining, Inference, and Prediction}, 2nd ed. New York, NY, USA: Springer, 2009.
+\vspace{-0.2cm}
 % sensitivity of DAS cables with slippage, straight or helical
-\bibitem{Kuvshinov2015} B.N. Kuvshinov, "Interaction of helically wound fibre-optic cables with plane seismic waves," \textit{Geophysical Prospecting}, vol. ***, no. ***, pp. 1-18, 2015.
-
+\bibitem{Kuvshinov2016} B.N. Kuvshinov, "Interaction of helically wound fibre-optic cables with plane seismic waves," \textit{Geophysical Prospecting}, vol. 64, no. 3, pp. 671-688, 2016.
+\vspace{-0.2cm}
+% review on clustering of time series data
+\bibitem{Liao2005} T.W. Liao, "Clustering of time series data- a survey," \textit{Pattern Recognition}, vol. 38, no. 11, pp. 1857-1874, 2005.
+\vspace{-0.2cm}
 % using horizontal components in interferometry as Rayleigh and Love wave components
@@
 J.
-Int.} vol.
+Int.}, vol.
 173, no.
@@
 281-298, 2008.
-
+\vspace{-0.2cm}
+% fundamental paper in ambient noise
+\bibitem{Lobkis2001} O.I. Lobkis and R.L. Weaver, "On the emergence of the Green's function in the correlations of a diffuse field," \textit{J. Acoust. Soc. Am.}, vol. 110, no. 6, pp. 3011-3017, 2001.
+\vspace{-0.2cm}
+% wavelet transforms and basic signal processing reference
+\bibitem{Mallat2008} S. Mallat, \textit{A Wavelet Tour of Signal Processing, Third Edition: The Sparse Way}. Burlington, MA, USA: Academic Press, 2008.
+\vspace{-0.2cm}
 % example of interferometry of ambient noise recorded by a trenched DAS array
@@
 2445-2450, 2015.
-
+\vspace{-0.2cm}
 % example of roadside DAS interferometry
@@
 2725-2629, 2016.
-
+\vspace{-0.2cm}
+% DAS conversion to velocities, and VSP example and side-by-side multi-mode, single-mode and 3C geophones
+\bibitem{Miller2016} D.E.
+Miller, T.M.
+Daley, B.M.
+Freifeld, M.
+Robertson, J.
+Cocker and M.
+Craven, "Simultaneous Acquisition of Distributed Acoustic Sensing VSP with Multi-mode and Single-mode FIber-optic Cables and 3C-Geophones at the Aquistore CO$_2$ Storage Site," \textit{CSEG Recorder}, pp.
+28-***, June 2016.
+\vspace{-0.2cm}
 % cross correlation, coherence and deconvolution, and using traffic noise
@@
 SA97-SA106, 2011.
-
+\vspace{-0.2cm}
 % ambient noise on dense array in Long Beach
@@
 1159-1173, 2015.
-
+\vspace{-0.2cm}
 % how DAS works
@@
 1688-1689, 2000.
-
+\vspace{-0.2cm}
 % overlapping windows, method of measuring convergence with correlation coefficient
 \bibitem{Seats2012} K.J. Seats, J.F. Lawrence and G.A. Prieto, "Improved ambient noise correlation functions using Welch's method," \textit{Geophys. J. Int.}, vol. 188, no. ***, pp. 513-523, 2012.
-
-% fundamental paper in ambient noise
-\bibitem{Lobkis2001} O.I. Lobkis and R.L. Weaver, "On the emergence of the Green's function in the correlations of a diffuse field," \textit{J. Acoust. Soc. Am.}, vol. 110, no. 6, pp. 3011-3017, 2001.
-
-% DAS conversion to velocities, and VSP example and side-by-side multi-mode, single-mode and 3C geophones
-\bibitem{Miller2016} D.E.
-Miller, T.M.
-Daley, B.M.
-Freifeld, M.
-Robertson, J.
-Cocker and M.
-Craven, "Simultaneous Acquisition of Distributed Acoustic Sensing VSP with Multi-mode and Single-mode FIber-optic Cables and 3C-Geophones at the Aquistore CO$_2$ Storage Site," \textit{CSEG Recorder}, pp.
-28-***, June 2016.
-
+\vspace{-0.2cm}
 % basics of interferometry
@@
 Campman and A.
-Verdel, "Tutorial on seismic interferometry: Part 1 - Basic principles and applications," \textit{Geophysics}, vol.
+Verdel, "Tutorial on seismic interferometry: Part 1- Basic principles and applications," \textit{Geophysics}, vol.
 75, no.
@@
 75A195-75A209, 2010.
+\vspace{-0.6cm}
 % more basics of interferometry
@@
 75A211-75A227, 2010.
-
+\vspace{-0.2cm}
 % example of DAS for microseismic event detection
@@
 2459-2463, 2013.
-
+\vspace{-0.2cm}
 % another example of DAS interferometry, this one side-by-side with geophones
@@
 Fratta, E.
-Matzel, PoroTomo Team, "High-resolution Shallow Structure Revealed with Ambient Noise Tomography on a Dense Array," \textit{Proceedings, 42nd Workshop on Geothermal Reservoir Engineering}, Stanford University, Stanford, CA, Feb.
+Matzel, PoroTomo Team, "High-resolution Shallow Structure Revealed with Ambient Noise Tomography on a Dense Array," \textit{Proceedings, 42nd Workshop on Geothermal Reservoir Engineering}, Stanford University, CA, Feb.
 13-15, 2017, SGP-TR-212.
-
+\vspace{-0.2cm}
 % velocity artifacts from ambient noise source changes
@@
 \end{thebibliography}
+\endgroup
 

### 2017-02-17 paper.tex
Update paper.tex
a little more motivation

@@
 
-\title{\vspace{-1.8cm} \Large Demystifying Ambient Seismic Noise Processing: \\Automating source analysis and processing decisions in the presence of coherent anthropogenic, natural, and sensor noise \vspace{-0.4cm}}
+\title{\vspace{-1.8cm} \Large Demystifying Ambient Seismic Noise Processing: \\Automating source analysis and processing decisions in the presence of coherent anthropogenic, natural, and sensor noise \vspace{-0.5cm}}
 \author[1,2]{\small Eileen R. Martin}
@@
 \author[3]{Martin Karrenbach}
-\author[1,2]{Biondo L. Biondi \vspace{-0.3cm}}
+\author[1,2]{Biondo L. Biondi \vspace{-0.4cm}}
 \affil[1]{Institute for Computational and Mathematical Engineering, Stanford University, Stanford, CA}
@@
 ****motivation for near surface seismic imaging****
-\\
-\\
-****motivation for using DAS****history \& current issues in DAS****
-Distributed acoustic sensing (DAS) probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors.
-It does this by performing optical interferometry on Rayleigh backscattered photons, associating the resulting phase shift with a subset of cable based on the two-way travel time of those photons, then applying a linear transformation to the change in phase shift associated with that subset of cable \cite{Posey2000}, \cite{Daley2013}, \cite{Bakku2015}.
-That cable can be either single- or multi-mode \cite{Miller2016}.
-For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber inside the cable to make its angular sensitivity more ideal for a variety of source-receiver geometries \cite{Kuvshinov2016}.
-\\
-\\
+\par
+Distributed acoustic sensing (DAS) probes a fiber optic cable (single or multi-mode \cite{Miller2016}) with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors.
+For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber in a helix \cite{Kuvshinov2016}.
+DAS is being increasingly adopted in the energy industry for microseismic monitoring \cite{Webster2013} and time-lapse seismic surveys \cite{Daley2013}, \cite{Mateeva2014}, \cite{Bakku2015}, \cite{Miller2016} because it allows for a dense series of seismic sensors to communicate over long distances while they are run by a single power source, and those sensors can be permanently left in the same location.
+\par
 ***motivation for using ambient noise****current issues in ambient noise*****
-\\
-\\
+\par
 There have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}.
-This type of installation allows ****
-By running fibers in existing telecommunications conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor are becoming a reality.
+In fact, we can sacrifice some ground-to-sensor coupling for easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}.
+By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor are becoming a reality.
 When the interrogator unit is probing a section of fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes.
+\par
+With large amounts of easily, cheaply recorded seismic data, we can only extract the full value of the data if we further automate the processing workflow.
 
@@
 \begin{itemize}
-\item overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with ambient noise in the presence of non-ideal noises
+\item overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with non-ideal ambient noise
 \vspace{-0.2cm}
@@
 \vspace{-0.2cm}
-It is difficult to gain intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics loosely laying in existing telecommunications conduits undernear the Stanford University campus.
-This particular case study is an ideal sandbox for a tutorial that demonstrates the wide variety of issues on the near horizon as we push for broader application of ambient noise studies.
-\\
-\\
+\par 
+It is hard to build intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics lying loosely in existing telecommunications conduits undernear the Stanford University campus.
+This particular experiment is ideal because it demonstrates the wide variety of issues on the near horizon as we push for broader use of ambient noise.
+\par
 We will discuss the noise level and sensitivities particular to recording with DAS, and a novel deployment that can greatly reduce costs (increasing data volumes): fiber loosely laying in existing telecommunications conduits.
@@
 	\vspace{-0.2cm}
-	\item Introduction: why we care about near surface imaging, discuss tradeoffs of sensor quality with cost and array density (DAS or MEMS, etc...), introduce idea of ambient noise to get rid of sources, ***
-	\vspace{-0.2cm}
+	\item Introduction: (i) why care about near surface imaging, (ii) tradeoffs of sensor quality with cost and array density (DAS or MEMS \cite{Evans2014}, etc...), (iii) introduce idea of ambient noise to get rid of active sources, (iv) today dense data acquisition over longer periods of time (potentially continuous streaming) is becoming easy, so we must further automate ambient noise processing
+	\vspace{-0.6cm}
 	\item Related work
@@
 		\vspace{-0.2cm}
-		\item Background on ambient noise: how it works that for the cost of just laying down receivers, you can extract signals mimicking those of an active seismic survey with a source at any receiver ****add citations******
-		\vspace{-0.2cm}
-		\item Background on Distributed Acoustic Sensing: how it works ****add citations****, how it has been used so far ****add citations****
-		\vspace{-0.2cm}
-		\item Examples of ambient seismic noise studies using Distributed Acoustic Sensing *****add citations****
+		\item Background on ambient noise: how you can extract signals mimicking those of an active seismic survey with a source at any receiver \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}, how non-ideal noise sources can cause artifacts in extracted velocities, and with careful processing choices this can be done in the presence of non-ideal noise sources \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even those occuring at higher frequencies \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}
+		\vspace{-0.1cm}
+		\item Background on Distributed Acoustic Sensing: how it works \cite{Posey2000}, \cite{Bakku2015}, how it has been used so far \cite{Daley2013}, \cite{Webster2013}, \cite{Mateeva2014}, and its limitations \cite{Kuvshinov2016}
+		\vspace{-0.1cm}
+		\item Examples of ambient noise studies with DAS \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Martin2017}, \cite{Zeng2017}
+		\vspace{-0.1cm}
+		\item Examples of how machine learning techniques have improved accuracy and sped up processing in other passive seismic applications \cite{Fisher2016}, \cite{Yoon2015}
 		\end{enumerate}
-	\vspace{-0.2cm}
-	\item Data Processing: we process the noise we record so that it mimicks data that would be recorded if the noise field followed theoretical ideals.
-Some common strategies include ****citations to review like papers*****
+	\vspace{-0.3cm}
+	\item Data Processing: we process the noise we record so that it mimicks data that would be recorded if the noise field followed theoretical ideals \cite{Bensen2007}
 		\begin{enumerate}
 		\vspace{-0.2cm}
-		\item Clustering/separation of data (k-means, hierarchical agglomerative, support vector machine separation) to speed up noise field exploration, identification of potential issues
+		\item New: clustering/separation of data (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009} to speed up noise field exploration, identification of potential issues
 		\vspace{-0.2cm}
-		\item How to use earthquakes to estimate sensor coupling heterogeneity when array installation is relatively uncontrolled (specifically, a fiber loosely laying in a variety of conduits) 
+		\item Existing tool: STA/LTA \cite{Withers1998} in time and Fourier domain to detect earthquakes or quarry blasts \cite{Bensen2007}, \cite{Girard2016}, this could potentially be enhanced to detect small events efficiently with template matching and statistical methods \cite{Yoon2015}
 		\vspace{-0.2cm}
-		\item Automatically detect \& filter persistent noise sources from short-time Fourier transforms
+		\item New: How to use earthquakes to estimate sensor coupling heterogeneity when array installation is relatively uncontrolled (loose fiber laying in a variety of conduits) 
 		\vspace{-0.2cm}
-		\item Automating the design of f-k domain filters for particular clusters based on separation from other clusters or background level noise.
+		\item New tool: Automatically detect \& filter persistent noise sources from short time spectra
+		\vspace{-0.6cm}
+		\item New: Automate the design of f-k domain filters for particular user-selected clusters based on separation from other clusters.
 In particular, a demonstration of removing vehicle noise: time domain mute vs.
@@
 automated design.
+		\vspace{-0.2cm}
+		\item New: In the presence of strong sources, cross-coherence (whitening after correlation) has a tendency to work better than cross-correlation in the sense of producing virtual source response estimates with fewer artifacts \cite{Nakata2011}, \cite{Martin2016}.
+Show metrics to automate choice of cross-correlation or cross-coherence per receiver pair (can change over time).
 		\end{enumerate}
 	\vspace{-0.2cm}
-	\item Measuring improvements: measure convergence rates of virtual source response function estimates like \cite{Seats2012}, but  (both cross-correlation and cross-coherence) for traditional processing and new processing methods
+	\item Measuring improvements: to compare old and new processing methods measure convergence of virtual source response function estimates (both cross-coherence and cross-correlation) as more data is incorporated using the correlation coefficient like \cite{Seats2012}.
+However, rather than using the first x amount of data, use a random sampling scheme over all the data so convergence rates get error bars.
 	\vspace{-0.2cm}
@@
 %***100 Eileen ***
-\vspace{-0.3cm}
+\vspace{-0.4cm}
 \subsection*{Biographies}
+\small
 \vspace{-0.2cm}
@@
 \\
-\textbf{Fantine Huot} \texttt{fantine@sep.stanford.edu} received her M.S.
+\textbf{Fantine Huot} (\texttt{fantine@sep.stanford.edu}) received her M.S.
 in Science and Engineering from the Ecole Normale Sup\'erieure des Mines de Paris in 2013.
@@
 \\
-\textbf{Biondo L. Biondi} \texttt{**email address} ****
+\textbf{Biondo L. Biondi} (\texttt{**email address}) ****
 
@@
 1239-1260, 2007.
-\vspace{-0.2cm}
-% overview of the array, reasons to use DAS in existing telecomm conduits
-\bibitem{Martin2017} E.R.
-Martin, B.L.
-Biondi, M.
-Karrenbach, S.
-Cole, "Continuous Subsurface Monitoring by Passice Seismic with Distributed Acoustic Sensors- The "Stanford Array" Experiment," \textit{Proceedings of the First EAGE Workshop on Practical Reservoir Monitoring}, 2017.
 \vspace{-0.2cm}
@@
 \vspace{-0.2cm}
+% overview of the array, reasons to use DAS in existing telecomm conduits
+\bibitem{Martin2017} E.R.
+Martin, B.L.
+Biondi, M.
+Karrenbach, S.
+Cole, "Continuous Subsurface Monitoring by Passice Seismic with Distributed Acoustic Sensors- The "Stanford Array" Experiment," \textit{Proceedings of the First EAGE Workshop on Practical Reservoir Monitoring}, 2017.
+\vspace{-0.2cm}
+\bibitem{Mateeva2014} A.
+Mateeva, J.
+Lopez, H.
+Potters, J.
+Mestayer, B.
+Cox, D.
+Kiyashchenko, P.
+Wills, S.
+Grandi, K.
+Hornman, B.
+Kuvshinov, W.
+Berlang, Z.
+Yang, R.
+Detomo, "Distributed acoustic sensing for reservoir monitoring with vertical seismic profiling," \textit{Geophysical Prospecting}, vol.
+62, no.
+4, pp.
+679-692, 2014.
+\vspace{-0.2cm}
 % DAS conversion to velocities, and VSP example and side-by-side multi-mode, single-mode and 3C geophones
@@
 \vspace{-0.2cm}
+% STA/LTA for trigger algorithms
+\bibitem{Withers1998} M.
+Withers, R.
+Aster, C.
+Young, J.
+Beiriger, M.
+Harris, S.
+Moore and J.
+Trujillo, "A comparison of select trigger algorithms for automated global seismic phase and event detection," \textit{B.
+Seismol.
+Soc.
+Am.}, vol.
+88, no.
+****, pp.
+95-106, 1998.
+\vspace{-0.2cm}
+% earthquake detection with templat ematching and machine learning
+\bibitem{Yoon2015} C.E.
+Yoon, O.
+O'Reilly, K.J.
+Bergen and G.C.
+Beroza, "Earthquake detection through computationally efficient similarity search," \textit{Science Advances}, vol.
+1, no.
+****, pp.
+e1501057-e1501057, 2015.
+\vspace{-0.2cm}
 % another example of DAS interferometry, this one side-by-side with geophones

### 2017-02-17 paper.tex
finished first draft

@@
 \vspace{-0.2cm}
-****motivation for near surface seismic imaging****
+By measuring the speed of seismic waves propagating in the Earth's near surface, we can image the top tens to hundreds of meters of the subsurface, with deeper features being resolved by lower frequencies.
+These seismic velocity images can be interpreted to evaluate earthquake or landslide risk, to detect permafrost, to find sinkholes or tunnels, or to track near surface changes related to drilling activities.
+Additionally, in cases of complex near-surface conditions, accurate near-surface images can be used to improve deeper seismic images.
 \par
-Distributed acoustic sensing (DAS) probes a fiber optic cable (single or multi-mode \cite{Miller2016}) with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors.
+By cross-correlating noise recorded at any two receivers in an array, we can extract signals mimicking an active seismic survey with a source at any receiver, virtual source response functions \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}.
+Thus, when active sources are too costly or logistically prohibitive, passive seismic can be a good option for near surface imaging.
+However, the theory is limited by the assumption of homogeneous uncorrelated sources \cite{Wapenaar2010A}.
+Non-ideal sources can cause artifacts in extracted velocities, but with careful processing experts can overcome these limitations \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even for anthropogenic sources occurring at higher frequencies \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}.
+Acquisition is a further issue: nearly all of the past ambient noise studies on dense arrays have been temporary arrays which were labor-intensive to install and maintain \cite{Lin2008}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}.
+\par
+Distributed acoustic sensing (DAS) probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors.
 For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber in a helix \cite{Kuvshinov2016}.
@@
 \par
-***motivation for using ambient noise****current issues in ambient noise*****
+The maintenance cost of permanent node arrays motivated the use of buried DAS, and there have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}.
+In fact, we can sacrifice some ground-to-sensor coupling in favor of easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}.
+By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor will soon be a reality.
+When the interrogator unit probes a fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes.
 \par
-There have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}.
-In fact, we can sacrifice some ground-to-sensor coupling for easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}.
-By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor are becoming a reality.
-When the interrogator unit is probing a section of fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes.
-\par
-With large amounts of easily, cheaply recorded seismic data, we can only extract the full value of the data if we further automate the processing workflow.
+With ambient noise data becoming increasingly easy to record, we can only extract its full value if we further automate the processing workflow.
 
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \subsection*{Significance} 
 \vspace{-0.2cm}
-This manuscript will walk through the basics (and some new tools) for processing ambient noise to do near-surface seismic imaging. Its purposes:
-\vspace{-0.2cm}
-\begin{itemize}
-\item overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with non-ideal ambient noise
-\vspace{-0.2cm}
-\item show optical engineers an application where an optical measurement device's noise level and sensitivity should be improved
-\vspace{-0.2cm}
-\item expose gepohysicists to an emerging acquisition method and a few machine learning tools and metrics to quantify processing decisions, a task that will become increasingly more difficult as data acquisition continues to become easier and cheaper
-\end{itemize}
-\vspace{-0.2cm}
+This manuscript will walk through the basics and some new tools for processing ambient noise for near-surface seismic imaging.
+Its purposes: (1) overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with non-ideal ambient noise, (2) show optical engineers an application where an optical measurement device's noise level and sensitivity should be improved, and (3) expose gepohysicists to an emerging acquisition method and a few machine learning tools and metrics to quantify processing decisions, a task that will become increasingly more difficult as data acquisition becomes easier and cheaper.
 \par 
-It is hard to build intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics lying loosely in existing telecommunications conduits undernear the Stanford University campus.
+It is hard to build intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics lying loosely in existing telecommunications conduits underneath the Stanford University campus.
 This particular experiment is ideal because it demonstrates the wide variety of issues on the near horizon as we push for broader use of ambient noise.
@@
 
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \subsection*{Outline of the paper}
@@
 	\item Abstract/summary
-	\vspace{-0.2cm}
+	\vspace{-0.3cm}
 	\item Introduction: (i) why care about near surface imaging, (ii) tradeoffs of sensor quality with cost and array density (DAS or MEMS \cite{Evans2014}, etc...), (iii) introduce idea of ambient noise to get rid of active sources, (iv) today dense data acquisition over longer periods of time (potentially continuous streaming) is becoming easy, so we must further automate ambient noise processing
-	\vspace{-0.6cm}
+	\vspace{-0.8cm}
 	\item Related work
 		\begin{enumerate}
-		\vspace{-0.2cm}
+		\vspace{-0.3cm}
 		\item Background on ambient noise: how you can extract signals mimicking those of an active seismic survey with a source at any receiver \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}, how non-ideal noise sources can cause artifacts in extracted velocities, and with careful processing choices this can be done in the presence of non-ideal noise sources \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even those occuring at higher frequencies \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}
@@
 		\end{enumerate}
-	\vspace{-0.3cm}
+	\vspace{-0.4cm}
 	\item Data Processing: we process the noise we record so that it mimicks data that would be recorded if the noise field followed theoretical ideals \cite{Bensen2007}
 		\begin{enumerate}
-		\vspace{-0.2cm}
+		\vspace{-0.3cm}
 		\item New: clustering/separation of data (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009} to speed up noise field exploration, identification of potential issues
-		\vspace{-0.2cm}
+		\vspace{-0.1cm}
 		\item Existing tool: STA/LTA \cite{Withers1998} in time and Fourier domain to detect earthquakes or quarry blasts \cite{Bensen2007}, \cite{Girard2016}, this could potentially be enhanced to detect small events efficiently with template matching and statistical methods \cite{Yoon2015}
-		\vspace{-0.2cm}
+		\vspace{-0.1cm}
 		\item New: How to use earthquakes to estimate sensor coupling heterogeneity when array installation is relatively uncontrolled (loose fiber laying in a variety of conduits) 
-		\vspace{-0.2cm}
-		\item New tool: Automatically detect \& filter persistent noise sources from short time spectra
-		\vspace{-0.6cm}
+		\vspace{-0.1cm}
+		\item New: Automatically detect \& filter persistent noise sources from short time spectra
+		\vspace{-0.1cm}
 		\item New: Automate the design of f-k domain filters for particular user-selected clusters based on separation from other clusters.
@@
 automated design.
-		\vspace{-0.2cm}
+		\vspace{-0.1cm}
 		\item New: In the presence of strong sources, cross-coherence (whitening after correlation) has a tendency to work better than cross-correlation in the sense of producing virtual source response estimates with fewer artifacts \cite{Nakata2011}, \cite{Martin2016}.
@@
 		\end{enumerate}
-	\vspace{-0.2cm}
-	\item Measuring improvements: to compare old and new processing methods measure convergence of virtual source response function estimates (both cross-coherence and cross-correlation) as more data is incorporated using the correlation coefficient like \cite{Seats2012}.
-However, rather than using the first x amount of data, use a random sampling scheme over all the data so convergence rates get error bars.
-	\vspace{-0.2cm}
+	\vspace{-0.5cm}
+	\item Measuring improvements: to compare old and new processing methods measure convergence of virtual source response function estimates (both cross-coherence and cross-correlation) as more data is incorporated using the correlation coefficient like \cite{Seats2012}, but use a random sampling scheme over all data so convergence rates have error bars.
+	\vspace{-0.3cm}
 	\item Show a few dispersion images and a velocity model resulting from surface wave tomography, a couple of the possible uses of the virtual source response estimates
-	\vspace{-0.2cm}
+	\vspace{-0.3cm}
 	\item Conclusions and future directions
@@
 
-%\begin{abstract} 
-%*****abstract words here - 150-200 Eileen*****
-%\end{abstract}
-
-%\section*{Motivation}
-%****what's ambient noise used for? - 200-250 Eileen****
-
-%****why use DAS? brief overview of its use****
-
-%\section*{New Contributions}
-%****what's new that we're doing, how it's different from what's been done before - 150 Eileen, little Fantine \& Yinbin****
-
-%\section*{Background Theory}
-%****ambient noise background - 150-220 Eileen****
-
-%\section*{Experimental Overview}
-%****overview of the array, potential noise sources, ambient noise field - 200 Eileen****
-%****Figure : 2 panels: Installation + map****
-
-%\section*{Data Processing}
-%***overview of data processing workflow - 200-250 Eileen****
-%***Figure : stft plots at a few different spots***
-%***Mention whitening here***
-%***Mention earthqukes and quarry blasts here***
-
-%\subsection*{Estimating Sensor Coupling Response}
-%***talk about different responses from different sensors and how to compensate for these, use many earthquakes for relative responses****
-%***100 Eileen***
-%***Figure : Boxed earthquake recording + response versus channel***
-
-%% \subsection*{Localizing Persistent Noise Sources}
-%% ***talk about localizing these things (motors, traffic, etc....) and show a little beamforming in small 1D parts of the array bandpassed to these persistent frequency peaks***
-
-%% \subsection*{Removing Earthquakes and Quarry Blasts}
-%% ****talk about getting rid of big events lighting up whole array****
-
-%\subsection*{Removing Vehicle Noise}
-%****talk about clustering and comparing noise sources****
-%***500 Fantine \& Yinbin***
-%***Figures : Before/after car mute + clusters and projection on map***
-%***Little Eileen for sentence about why cars are a problem***
-
-%% \subsection*{Removing Correlated Noise Sources}
-%% *****talk about detecting temporally and spatially correlated noise sources, more generally than cars****
-
-%% \subsection*{Traveltime Picking}
-%% *** talk about traveltime picking before Rayleigh wave tomography****
-%% *** STANDBY ***
-
-%% \subsection*{Tomography}
-%% **** talk about how near-surface tomography is set up****
-%% *** STANDBY ***
-
-%\section*{Results}
-%***overview of results****
-%*** Convergence rates
-%***150 Eileen *** 
-%***Figure : line plot of correlation coeeficients vs epoch/time***
-
-%% \subsection*{Convergence Rate Improvement Due to Pre-Processing}
-%***show comparisons of cross-correlations with/without sensor coupling response, and with scaling by energy per channel of cross-correlations (also show spectrum per channel of these)****
-%\\
-%******show comparisons of convergence rates with/without vehicles removed, and with vehicles just removed via STA/LTA jumps****
-
-%% \subsection*{Tomography Along Array Lines}
-%% ***show tomography results along multiple lines on campus****
-%% *** STANDBY ***
-
-%\section*{Summary}
-%****summary words****
-%***100 Eileen ***
-\vspace{-0.4cm}
+\vspace{-0.9cm}
 \subsection*{Biographies}
@@
 student in the Geophysics Department.
-She first started working with ambient noise and DAS data from permafrost thaw monitoring experiments done by Lawrence Berkeley National Lab and the US Army Corps of Engineers.
+She is also affiliated with Lawrence Berkeley National Lab where she collaborates with a team on fiber optic monitoring of permafrost thaw and on the use of dark fiber for event detection.
 \\
@@
 in Geophysics at Stanford University and joined the Stanford Exploration Project.
-She works on addressing earth science problems by leveraging machine learning algorithms.
+She leverages machine learning algorithms to address earth science problems.
 \\
-\textbf{Yinbin Ma} \texttt{**email address} *****
+\textbf{Yinbin Ma} (\texttt{yinbin@stanford.edu}) is a Ph.D.
+student in the Institute for Computational and Mathematical Engineering at Stanford.
+He works on time-lapse seismic imaging at the Stanford Exploration Project.
 \\
@@
 
-\vspace{-0.3cm}
+\vspace{-0.6cm}
 \begingroup
@@
 \begin{thebibliography}{30} % no more than 30 references
-\vspace{-0.2cm}
+\vspace{-0.3cm}
 \small
@@
 % review on clustering of time series data
-\bibitem{Liao2005} T.W. Liao, "Clustering of time series data- a survey," \textit{Pattern Recognition}, vol. 38, no. 11, pp. 1857-1874, 2005.
+\bibitem{Liao2005} T.W. Liao, "Clustering of time series data- a survey," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.
 \vspace{-0.2cm}

### 2017-02-17 paper.tex
polishing up a few grammar errors and missing info in bibliography

@@
 \par
-By cross-correlating noise recorded at any two receivers in an array, we can extract signals mimicking an active seismic survey with a source at any receiver, virtual source response functions \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}.
+By cross-correlating noise recorded at any two receivers in an array, we can extract signals mimicking an active seismic survey with a source at any receiver, called virtual source response functions \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}.
 Thus, when active sources are too costly or logistically prohibitive, passive seismic can be a good option for near surface imaging.
@@
 		\vspace{-0.3cm}
-		\item New: clustering/separation of data (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009} to speed up noise field exploration, identification of potential issues
+		\item New: clustering/separation of data with features including wavelet transforms (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009}, \cite{Mallat2008} to speed up noise field exploration, identification of potential issues
 		\vspace{-0.1cm}
@@
 Ulrich and A.
-Wagner, "A Field Test of Distributed Acoustic Sensing for Ambient Noise Recording," \textit{SEG Technical Program Expanded Abstracts}, pp.
+Wagner, ``A Field Test of Distributed Acoustic Sensing for Ambient Noise Recording," \textit{SEG Technical Program Expanded Abstracts}, pp.
 2620-2624, 2015.
@@
 % working with DAS data, how DAS works basically
-\bibitem{Bakku2015} S.K. Bakku, "Fracture Characterization from Seismic Measurements in a Borehole," Ph.D. Thesis, Massachusetts Institute of Technology, 2015.
+\bibitem{Bakku2015} S.K. Bakku, ``Fracture Characterization from Seismic Measurements in a Borehole," Ph.D. Thesis, Massachusetts Institute of Technology, 2015.
 \vspace{-0.2cm}
@@
 Shapiro and Y.
-Yang, "Processing seismic ambient noise data to obtain reliable broad-band surface wave dispersion measurements," \textit{Geophys.
+Yang, ``Processing seismic ambient noise data to obtain reliable broad-band surface wave dispersion measurements," \textit{Geophys.
 J.
@@
 169, no.
-****, pp.
+3, pp.
 1239-1260, 2007.
@@
 Papanicolaou and C.
-Tsoga, "Robust seismic velocity change estimation using ambient noise recordings," \textit{Geophys.
+Tsoga, ``Robust seismic velocity change estimation using ambient noise recordings," \textit{Geophys.
 J.
@@
 Ajo-Franklin and S.
-Dou, "Field testing of fiber-optic distributed acoustic sensing (DAS) for subsurface seismic monitoring," \textit{The Leading Edge}, pp.
+Dou, ``Field testing of fiber-optic distributed acoustic sensing (DAS) for subsurface seismic monitoring," \textit{The Leading Edge}, pp.
 936-942, June 2013.
@@
 Hellweg and J.F.
-Lawrence, "Performance of Several Low-Cost Accelerometers," \textit{Seismol.
+Lawrence, ``Performance of Several Low-Cost Accelerometers," \textit{Seismol.
 Res.
@@
 Camp and V.V.
-Krzhizhanovskaya, "Anomaly detection in earth dam and levee passive seismic data using support vector machines and automatic feature selection," \textit{J.
+Krzhizhanovskaya, ``Anomaly detection in earth dam and levee passive seismic data using support vector machines and automatic feature selection," \textit{J.
 Comput.
@@
 % ambient noise with mine blasts (similar to earthquakes)
-\bibitem{Girard2016} A.J. Girard and J. Shragge, "Extracting body waves from ambient seismic recordings," \textit{SEG Technical Program Expanded Abstracts}, pp. 2715-2719, 2016.
+\bibitem{Girard2016} A.J. Girard and J. Shragge, ``Extracting body waves from ambient seismic recordings," \textit{SEG Technical Program Expanded Abstracts}, pp. 2715-2719, 2016.
 \vspace{-0.2cm}
@@
 % sensitivity of DAS cables with slippage, straight or helical
-\bibitem{Kuvshinov2016} B.N. Kuvshinov, "Interaction of helically wound fibre-optic cables with plane seismic waves," \textit{Geophysical Prospecting}, vol. 64, no. 3, pp. 671-688, 2016.
+\bibitem{Kuvshinov2016} B.N. Kuvshinov, ``Interaction of helically wound fibre-optic cables with plane seismic waves," \textit{Geophysical Prospecting}, vol. 64, no. 3, pp. 671-688, 2016.
 \vspace{-0.2cm}
 % review on clustering of time series data
-\bibitem{Liao2005} T.W. Liao, "Clustering of time series data- a survey," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.
+\bibitem{Liao2005} T.W. Liao, ``Clustering of time series data- a survey," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.
 \vspace{-0.2cm}
@@
 Moschetti and M.H.
-Ritzwoller, "Surface wave tomography of the western United States from ambient seismic noise: Rayleigh and Love wave phase velocity maps," \textit{Geophys.
+Ritzwoller, ``Surface wave tomography of the western United States from ambient seismic noise: Rayleigh and Love wave phase velocity maps," \textit{Geophys.
 J.
@@
 173, no.
-***, pp.
+1, pp.
 281-298, 2008.
@@
 % fundamental paper in ambient noise
-\bibitem{Lobkis2001} O.I. Lobkis and R.L. Weaver, "On the emergence of the Green's function in the correlations of a diffuse field," \textit{J. Acoust. Soc. Am.}, vol. 110, no. 6, pp. 3011-3017, 2001.
+\bibitem{Lobkis2001} O.I.
+Lobkis and R.L.
+Weaver, ``On the emergence of the Green's function in the correlations of a diffuse field," \textit{J.
+Acoust.
+Soc.
+Am.}, vol.
+110, no.
+6, pp.
+3011-3017, 2001.
 \vspace{-0.2cm}
@@
 Wagner, C.
-Ulrich, "Interferometry of ambient noise from a trenched distributed acoustic sensing array," \textit{SEG Technical Program Expanded Abstracts}, pp.
+Ulrich, ``Interferometry of ambient noise from a trenched distributed acoustic sensing array," \textit{SEG Technical Program Expanded Abstracts}, pp.
 2445-2450, 2015.
@@
 Wagner and K.
-Bjella, "Interferometry of a roadside DAS array in Fairbanks, AK," \textit{SEG Technical Program Expanded Abstracts}, pp.
+Bjella, ``Interferometry of a roadside DAS array in Fairbanks, AK," \textit{SEG Technical Program Expanded Abstracts}, pp.
 2725-2629, 2016.
@@
 Karrenbach, S.
-Cole, "Continuous Subsurface Monitoring by Passice Seismic with Distributed Acoustic Sensors- The "Stanford Array" Experiment," \textit{Proceedings of the First EAGE Workshop on Practical Reservoir Monitoring}, 2017.
+Cole, ``Continuous Subsurface Monitoring by Passice Seismic with Distributed Acoustic Sensors- The `Stanford Array' Experiment," \textit{Proceedings of the First EAGE Workshop on Practical Reservoir Monitoring}, 2017.
 \vspace{-0.2cm}
@@
 Yang, R.
-Detomo, "Distributed acoustic sensing for reservoir monitoring with vertical seismic profiling," \textit{Geophysical Prospecting}, vol.
+Detomo, ``Distributed acoustic sensing for reservoir monitoring with vertical seismic profiling," \textit{Geophysical Prospecting}, vol.
 62, no.
@@
 Cocker and M.
-Craven, "Simultaneous Acquisition of Distributed Acoustic Sensing VSP with Multi-mode and Single-mode FIber-optic Cables and 3C-Geophones at the Aquistore CO$_2$ Storage Site," \textit{CSEG Recorder}, pp.
-28-***, June 2016.
+Craven, ``Simultaneous Acquisition of Distributed Acoustic Sensing VSP with Multi-mode and Single-mode FIber-optic Cables and 3C-Geophones at the Aquistore CO$_2$ Storage Site," \textit{CSEG Recorder}, vol.
+41, no.
+6, pp.
+28-33, 2016.
 \vspace{-0.2cm}
@@
 Larner and T.
-Matsuoka, "Shear wave imaging from traffic noise using seismic interferometry by cross-coherence," \textit{Geophysics}, vol.
+Matsuoka, ``Shear wave imaging from traffic noise using seismic interferometry by cross-coherence," \textit{Geophysics}, vol.
 76, no.
@@
 Lawrence and P.
-Bou\'{e}, "Body wave extraction and tomography at Long Beach, California, with ambient-noise interferometry," \textit{J.
+Bou\'{e}, ``Body wave extraction and tomography at Long Beach, California, with ambient-noise interferometry," \textit{J.
 Geophys.
@@
 Johnson and S.T.
-Vohra, "Strain sensing based on coherence Rayleigh scattering in an optical fibre," \textit{Electronics Letters}, vol.
+Vohra, ``Strain sensing based on coherence Rayleigh scattering in an optical fibre," \textit{Electronics Letters}, vol.
 36, no.
@@
 % overlapping windows, method of measuring convergence with correlation coefficient
-\bibitem{Seats2012} K.J. Seats, J.F. Lawrence and G.A. Prieto, "Improved ambient noise correlation functions using Welch's method," \textit{Geophys. J. Int.}, vol. 188, no. ***, pp. 513-523, 2012.
+\bibitem{Seats2012} K.J. Seats, J.F. Lawrence and G.A. Prieto, ``Improved ambient noise correlation functions using Welch's method," \textit{Geophys. J. Int.}, vol. 188, no. 2, pp. 513-523, 2012.
 \vspace{-0.2cm}
@@
 Campman and A.
-Verdel, "Tutorial on seismic interferometry: Part 1- Basic principles and applications," \textit{Geophysics}, vol.
+Verdel, ``Tutorial on seismic interferometry: Part 1- Basic principles and applications," \textit{Geophysics}, vol.
 75, no.
@@
 Snieder and A.
-Curtis, "Tutorial on seismic interferometry: Part 2 - Underlying theory and new advances," \textit{Geophysics}, vol.
+Curtis, ``Tutorial on seismic interferometry: Part 2 - Underlying theory and new advances," \textit{Geophysics}, vol.
 75, no.
@@
 Perkins and M.
-Molenaar, "Micro-Seismic Detection using Distributed Acoustic Sensing," \textit{SEG Technical Program Expanded Abstracts}, pp.
+Molenaar, ``Micro-Seismic Detection using Distributed Acoustic Sensing," \textit{SEG Technical Program Expanded Abstracts}, pp.
 2459-2463, 2013.
@@
 Moore and J.
-Trujillo, "A comparison of select trigger algorithms for automated global seismic phase and event detection," \textit{B.
+Trujillo, ``A comparison of select trigger algorithms for automated global seismic phase and event detection," \textit{B.
 Seismol.
@@
 88, no.
-****, pp.
+1, pp.
 95-106, 1998.
@@
 Bergen and G.C.
-Beroza, "Earthquake detection through computationally efficient similarity search," \textit{Science Advances}, vol.
+Beroza, ``Earthquake detection through computationally efficient similarity search," \textit{Science Advances}, vol.
 1, no.
-****, pp.
-e1501057-e1501057, 2015.
+11, pp.
+e1501057, 2015.
 \vspace{-0.2cm}
@@
 Fratta, E.
-Matzel, PoroTomo Team, "High-resolution Shallow Structure Revealed with Ambient Noise Tomography on a Dense Array," \textit{Proceedings, 42nd Workshop on Geothermal Reservoir Engineering}, Stanford University, CA, Feb.
+Matzel, PoroTomo Team, ``High-resolution Shallow Structure Revealed with Ambient Noise Tomography on a Dense Array," \textit{Proceedings, 42nd Workshop on Geothermal Reservoir Engineering}, Stanford University, CA, Feb.
 13-15, 2017, SGP-TR-212.
@@
 Tsai and R.W.
-Clayton, "Spurious velocity changes caused by temporal variations in ambient noise frequency content," \textit{Geophys.
+Clayton, ``Spurious velocity changes caused by temporal variations in ambient noise frequency content," \textit{Geophys.
 J.
@@
 194, no.
-***, pp.
+3, pp.
 1574-1581, 2013.

### 2017-04-27 whitepaper.tex
added some files to modify asap

--- 2017-02-17 paper.tex
@@
 
-\title{\vspace{-1.8cm} \Large Demystifying Ambient Seismic Noise Processing: \\Automating source analysis and processing decisions in the presence of coherent anthropogenic, natural, and sensor noise \vspace{-0.5cm}}
+\title{\vspace{-1.8cm} \Large Demystifying Ambient Seismic Noise Analysis: \\Automating processing decisions in the presence of coherent noise \vspace{-0.5cm}}
 \author[1,2]{\small Eileen R. Martin}
@@
 \author[3]{Martin Karrenbach}
-\author[1,2]{Biondo L. Biondi \vspace{-0.4cm}}
+\author[2,1]{Biondo L. Biondi \vspace{-0.4cm}}
 \affil[1]{Institute for Computational and Mathematical Engineering, Stanford University, Stanford, CA}
@@
 \vspace{-0.2cm}
-By measuring the speed of seismic waves propagating in the Earth's near surface, we can image the top tens to hundreds of meters of the subsurface, with deeper features being resolved by lower frequencies.
+By measuring the speed of seismic waves propagating in the Earth's near-surface, we can image the top tens to hundreds of meters of the subsurface, with deeper features being resolved by lower frequencies.
 These seismic velocity images can be interpreted to evaluate earthquake or landslide risk, to detect permafrost, to find sinkholes or tunnels, or to track near surface changes related to drilling activities.
-Additionally, in cases of complex near-surface conditions, accurate near-surface images can be used to improve deeper seismic images.
+Additionally, in cases of complex near-surface conditions, resolving this complexity is a prerequisite to obtaining a high-quality image of the deeper subsurface.
 \par
-By cross-correlating noise recorded at any two receivers in an array, we can extract signals mimicking an active seismic survey with a source at any receiver, called virtual source response functions \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}.
-Thus, when active sources are too costly or logistically prohibitive, passive seismic can be a good option for near surface imaging.
+By cross-correlating noise recorded at a selected receiver with noise recorded by all other receivers in an array, we can extract signals mimicking an active seismic survey with a source at the selected receiver, called its virtual source response function \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}.
+Thus, when active sources are too costly or logistically prohibitive, passive seismic can be a good option for near-surface imaging.
 However, the theory is limited by the assumption of homogeneous uncorrelated sources \cite{Wapenaar2010A}.
@@
 \par
-Distributed acoustic sensing (DAS) probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors.
-For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber in a helix \cite{Kuvshinov2016}.
-DAS is being increasingly adopted in the energy industry for microseismic monitoring \cite{Webster2013} and time-lapse seismic surveys \cite{Daley2013}, \cite{Mateeva2014}, \cite{Bakku2015}, \cite{Miller2016} because it allows for a dense series of seismic sensors to communicate over long distances while they are run by a single power source, and those sensors can be permanently left in the same location.
+Distributed acoustic sensing (DAS) is a new acquisition technology being increasingly adopted in the energy industry for microseismic monitoring \cite{Webster2013} and time-lapse seismic surveys \cite{Daley2013}, \cite{Mateeva2014}, \cite{Bakku2015}, \cite{Miller2016}.
+DAS probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain sensors.
+Most experiments use low-cost straight fibers, but these fibers only measure strain along the axis of the fiber \cite{Kuvshinov2016}.
+This downside is often outweighed by the benefit of a dense series of permanently installed seismic sensors communicating over long distances and running on a single power source.
 \par
-The maintenance cost of permanent node arrays motivated the use of buried DAS, and there have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}.
+Motivated by avoiding the maintenance cost of node arrays, there have been several recent ambient noise experiments using fiber optics buried in shallow trenches directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}.
 In fact, we can sacrifice some ground-to-sensor coupling in favor of easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}.
-By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor will soon be a reality.
-When the interrogator unit probes a fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes.
+By running fibers in existing conduits, or even plugging into unused fibers in existing telecommunications bundles, easy, on-demand, repeatable seismic studies (even in urban areas) will soon be a reality.
+When the IU probes a fiber, the data can serve many purposes: near-surface imaging with ambient noise, earthquake detection \cite{Martin2017} and location, studying local site response to earthquakes, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes.
 \par
-With ambient noise data becoming increasingly easy to record, we can only extract its full value if we further automate the processing workflow.
+With ambient noise data becoming increasingly easy to record, data volumes are increasing, and we can only extract their full value if we further automate the processing workflow.
 
@@
 \vspace{-0.2cm}
-This manuscript will walk through the basics and some new tools for processing ambient noise for near-surface seismic imaging.
-Its purposes: (1) overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with non-ideal ambient noise, (2) show optical engineers an application where an optical measurement device's noise level and sensitivity should be improved, and (3) expose gepohysicists to an emerging acquisition method and a few machine learning tools and metrics to quantify processing decisions, a task that will become increasingly more difficult as data acquisition becomes easier and cheaper.
+This manuscript will walk through the basics of processing ambient noise for near-surface imaging, and introduce some tools that are new to this application area, but already used for other spatial and temporal data.
+Its purposes: (1) overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with non-ideal ambient noise, (2) show optical engineers an application where an optical measurement device's noise level and sensitivity should be improved, and (3) expose geophysicists to an emerging acquisition method and a few machine learning tools and metrics to quantify processing decisions, a task that will become increasingly more difficult as data acquisition becomes easier and cheaper.
 \par 
-It is hard to build intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics lying loosely in existing telecommunications conduits underneath the Stanford University campus.
-This particular experiment is ideal because it demonstrates the wide variety of issues on the near horizon as we push for broader use of ambient noise.
+It is difficult to build intuition about ambient noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics lying loosely in existing telecommunications conduits underneath the Stanford University campus.
+This particular experiment is ideal because it demonstrates the wide variety of issues coming up on the horizon as we push for broader use of ambient noise.
 \par
-We will discuss the noise level and sensitivities particular to recording with DAS, and a novel deployment that can greatly reduce costs (increasing data volumes): fiber loosely laying in existing telecommunications conduits.
+We will discuss the noise and sensitivities particular to recording with DAS, and a novel deployment that can greatly reduce costs: fiber loosely laying in existing telecommunications conduits.
 The array detects a wide variety of seismic noise sources that do not conform to the ideals of existing ambient noise theory: it sits in a seismically active region, 20 km from the Pacific ocean, 7 km from the San Francisco bay, with major highways on either side, a variety of roads with differing levels of traffic near the fiber, regular quarry blasts within 15 km, plumbing and HVAC systems throughout the site, multiple construction sites near the array, and foot and bicycle traffic throughout.
-With over 600 sensors continuously recording 50 samples per second since September 2016, manual inspection of most data is infeasible, making the automation tools shown in this paper critical to extracting subsurface information from these data.
+With over 600 sensors continuously recording 50 samples per second since September 2016, manual inspection of most data is infeasible, making the automation tools shown in this paper critical to extracting subsurface information from the data.
 
@@
 	\vspace{-0.3cm}
-	\item Introduction: (i) why care about near surface imaging, (ii) tradeoffs of sensor quality with cost and array density (DAS or MEMS \cite{Evans2014}, etc...), (iii) introduce idea of ambient noise to get rid of active sources, (iv) today dense data acquisition over longer periods of time (potentially continuous streaming) is becoming easy, so we must further automate ambient noise processing
-	\vspace{-0.8cm}
-	\item Related work
+	\item Introduction: (i) use of near surface imaging, (ii) tradeoffs: sensor quality vs.
+cost and array density (DAS or MEMS \cite{Evans2014}, etc...), (iii) idea of ambient noise to get rid of active sources, (iv) dense data acquisition over longer periods of time (potentially continuously streaming in as it's recorded) is becoming easy, so we must further automate ambient noise processing.
+	\vspace{-0.3cm}
+	\item Related work:
 		\begin{enumerate}
 		\vspace{-0.3cm}
-		\item Background on ambient noise: how you can extract signals mimicking those of an active seismic survey with a source at any receiver \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}, how non-ideal noise sources can cause artifacts in extracted velocities, and with careful processing choices this can be done in the presence of non-ideal noise sources \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even those occuring at higher frequencies \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}
+		\item Background on ambient noise: how you can extract signals mimicking those of an active seismic survey with a source at any receiver \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}, how non-ideal noise sources can cause artifacts in extracted velocities, and with careful processing choices this can be done in the presence of non-ideal noise sources \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even higher frequencies sources \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}
 		\vspace{-0.1cm}
-		\item Background on Distributed Acoustic Sensing: how it works \cite{Posey2000}, \cite{Bakku2015}, how it has been used so far \cite{Daley2013}, \cite{Webster2013}, \cite{Mateeva2014}, and its limitations \cite{Kuvshinov2016}
+		\item Background on Distributed Acoustic Sensing: how it works \cite{Bakku2015}, \cite{Posey2000}, how it has been used so far \cite{Bakku2015}, \cite{Daley2013}, \cite{Webster2013}, \cite{Mateeva2014}, and its limitations \cite{Kuvshinov2016}
 		\vspace{-0.1cm}
@@
 		\vspace{-0.1cm}
-		\item Examples of how machine learning techniques have improved accuracy and sped up processing in other passive seismic applications \cite{Fisher2016}, \cite{Yoon2015}
+		\item Examples of how machine learning techniques have improved accuracy and sped up processing in other seismic applications \cite{Fisher2016}, \cite{Yoon2015}
 		\end{enumerate}
 	\vspace{-0.4cm}
-	\item Data Processing: we process the noise we record so that it mimicks data that would be recorded if the noise field followed theoretical ideals \cite{Bensen2007}
+	\item Data Processing on recorded noise so it mimics data from theoretically ideal noise field \cite{Bensen2007}
 		\begin{enumerate}
 		\vspace{-0.3cm}
-		\item New: clustering/separation of data with features including wavelet transforms (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009}, \cite{Mallat2008} to speed up noise field exploration, identification of potential issues
+		\item New: tools to aid people in speeding up noise field exploration and identification of potential issues including clustering/separation of data with features including wavelet transforms (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009}, \cite{Mallat2008}
 		\vspace{-0.1cm}
-		\item Existing tool: STA/LTA \cite{Withers1998} in time and Fourier domain to detect earthquakes or quarry blasts \cite{Bensen2007}, \cite{Girard2016}, this could potentially be enhanced to detect small events efficiently with template matching and statistical methods \cite{Yoon2015}
+		\item Common tool: STA/LTA \cite{Withers1998} in time and Fourier domain to find earthquakes or quarry blasts \cite{Bensen2007}, \cite{Girard2016}, may better detect small events efficiently with fingerprint search \cite{Yoon2015}
 		\vspace{-0.1cm}
-		\item New: How to use earthquakes to estimate sensor coupling heterogeneity when array installation is relatively uncontrolled (loose fiber laying in a variety of conduits) 
+		\item New: How to use earthquakes to estimate sensor coupling heterogeneity when array installation is relatively uncontrolled (loose fiber laying in variety of conduits) 
 		\vspace{-0.1cm}
@@
 		\vspace{-0.1cm}
-		\item New: In the presence of strong sources, cross-coherence (whitening after correlation) has a tendency to work better than cross-correlation in the sense of producing virtual source response estimates with fewer artifacts \cite{Nakata2011}, \cite{Martin2016}.
-Show metrics to automate choice of cross-correlation or cross-coherence per receiver pair (can change over time).
+		\item New: In presence of strong/correlated sources, cross-coherence tends to work better than cross-correlation in the sense of producing virtual source response estimates with fewer artifacts \cite{Nakata2011}, \cite{Martin2016}.
+Show metrics to automate choice of stack weighting of cross-correlation or cross-coherence per receiver pair (can change over time).
 		\end{enumerate}
 	\vspace{-0.5cm}
-	\item Measuring improvements: to compare old and new processing methods measure convergence of virtual source response function estimates (both cross-coherence and cross-correlation) as more data is incorporated using the correlation coefficient like \cite{Seats2012}, but use a random sampling scheme over all data so convergence rates have error bars.
+	\item Measuring improvements: to compare old and new processing methods, measure convergence of virtual source response function estimates (both cross-coherence and cross-correlation) as more data is incorporated using the correlation coefficient like \cite{Seats2012}, but use a random sampling scheme over all data so convergence rates have error bars.
 	\vspace{-0.3cm}
-	\item Show a few dispersion images and a velocity model resulting from surface wave tomography, a couple of the possible uses of the virtual source response estimates
+	\item Show a few dispersion images and a velocity model resulting from surface wave tomography, a couple of the possible uses of the virtual source response estimates.
 	\vspace{-0.3cm}
@@
 \\
-\textbf{Martin Karrenbach} (\texttt{***email address}) *******
+\textbf{Martin Karrenbach} (\texttt{Martin.Karrenbach@optasense.com}) Martin Karrenbach is the senior manager of innovation at OptaSense in Brea, CA.
+He received his Ph.D.
+in geophysics from Stanford University in 1995.
 \\
-\textbf{Biondo L. Biondi} (\texttt{**email address}) ****
+\textbf{Biondo L.
+Biondi} (\texttt{biondo@stanford.edu}) Biondo Biondi is a professor in the geophysics department at Stanford University, and leads the Stanford Exploration Project.
+He received his Ph.D.
+in geophysics from Stanford in 1990.
+He published the first book on 3D reflection seismology ("3D Seismic Imaging") and pioneered many methods in wavefield migration and velocity analysis.
 
@@
 % review on clustering of time series data
-\bibitem{Liao2005} T.W. Liao, ``Clustering of time series data- a survey," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.
+\bibitem{Liao2005} T.W. Liao, ``Clustering of time series data," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.
 \vspace{-0.2cm}
@@
 % wavelet transforms and basic signal processing reference
-\bibitem{Mallat2008} S. Mallat, \textit{A Wavelet Tour of Signal Processing, Third Edition: The Sparse Way}. Burlington, MA, USA: Academic Press, 2008.
+\bibitem{Mallat2008} S. Mallat, \textit{A Wavelet Tour of Signal Processing}, 3rd ed. Burlington, MA, USA: Academic Press, 2008.
 \vspace{-0.2cm}

### 2017-04-27 code/src/calc_oneBitXCorr.py

@@ -1,0 +1,158 @@
+#cimport numpy as np
+import numpy as np
+import matplotlib.pyplot as plt
+import datetime as dt
+import struct
+import obspy
+import sys
+import xcorr
+
+sys.path.append('/home/ermartin/PassiveSeismicArray')
+# get parameters: startTime, secondsPerFile, secondsPerWindowWidth, secondsPerWindowOffset, xCorrMaxTimeLagSeconds, nFiles, outfilePath, outfileList, srcChList, startCh, endCh, minFrq, maxFrq
+paramsPath = sys.argv[1]
+sys.path.append(paramsPath)
+from params import *
+import fileSet as fs
+import regularFileSet as rfs
+
+nTxtFileHeader = 3200
+nBinFileHeader = 400
+nTraceHeader = 240
+def readTrace(infile,nSamples,dataLen,traceNumber,endian,startSample,nSamplesToRead):
+    '''infile is .sgy, nSamples is the number of samples per sensor, and traceNumber is the sensor number (start with 1),dataLen is number of bytes per data sample'''
+
+    fin = open(infile, 'rb') # open file for reading binary mode
+    startData = nTxtFileHeader+nBinFileHeader+nTraceHeader+(traceNumber-1)*(nTraceHeader+dataLen*nSamples)+startSample*dataLen
+    fin.seek(startData)
+    thisTrace = np.zeros(nSamplesToRead)
+    for i in range(nSamplesToRead):
+       	# was >f before
+       	thisTrace[i] = struct.unpack(endian+'f',fin.read(dataLen))[0]
+    fin.close()
+    return thisTrace
+
+
+#cdef extern from "xcorrmodule.h":
+#	cdef cppclass 
+        
+def crossCorrOneBit(virtualSrcTrace, allOtherReceiversTraces, nLagSamples): #, version = 'C'):
+    #if(version == 'C'):
+	#xCorr = xcorr.oneBitXCorr(virtualSrcTrace, allOtherReceiversTraces, nLagSamples)
+    #else:
+    numberChannels = allOtherReceiversTraces.shape[0]
+    nt = allOtherReceiversTraces.shape[1]
+    xCorr = np.empty((numberChannels,1+2*nLagSamples),dtype=np.int32)
+    sumWidth = nt-2*nLagSamples
+    for i in range(-1*nLagSamples,nLagSamples+1):
+        startSample = i + nLagSamples
+        endSample = startSample + sumWidth
[…]

### 2017-04-27 code/src/corrsmodule.cpp

@@ -3,13 +3,14 @@
 #include <cmath>
 #include <iostream>
 #include <vector>
+#include <cfloat>
 
 float _correlationCoeff(float *longTermFct, float *shortTermFct){
   /// Calculates the zero time lag correlation 
   /// but does so with blocked sums (so no overflow)
 
   float corrCoeff = 1.0;
-  /*float partialSum = 0.0;
+  float partialSum = 0.0;
   int len = ****;
 
   // break arrays into blocks
@@ -21,7 +22,7 @@ float _correlationCoeff(float *longTermFct, float *shortTermFct){
     }
   }
   // add any stragler terms at the end
-  corrCoeff += double(partialSum);*/
+  corrCoeff += double(partialSum);
 
   return corrCoeff;
 }

### 2017-04-27 code/src/params/params1/params.py

@@ -1,0 +1,15 @@
+import datetime as dt
+
+startTime = dt.datetime(2016,9,4,5,0,54,932*1000)
+secondsPerFile = 60
+secondsPerWindowWidth = 300
+secondsPerWindowOffset = 150
+xCorrMaxTimeLagSeconds = 3.0
+nFiles = 240
+outfilePath = '/scratch/ermartin/oneBitXCorr/'
+outfileListFile = '/home/ermartin/StanfordDASArray/inter_results/outfileList1.txt'
+srcChList = 'sourceList.txt'
+startCh = 15
+endCh = 300
+minFrq = 0.2
+maxFrq = 24.0

### 2017-04-27 code/src/plot_oneBitXCorr.py

@@ -1,0 +1,67 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+paramsPath = sys.argv[1]
+sys.path.append(paramsPath)
+from params import *
+
+srcFile = open(srcChList,'r')
+srcChannelsStrings = srcFile.readlines()
+srcFile.close()
+srcChannels = [int(ch) for ch in srcChannelsStrings]
+
+# get the list of all the files
+outFileList = open(outfileListFile,'r')
+listOfAllFiles = outFileList.readlines()
+listOfAllFiles = [(f.strip('\n'))+'.npz' for f in listOfAllFiles]
+outFileList.close()
+
+for ch in srcChannels:
+    # get the list of file names for this source channel
+    containsPhrase = '_srcCh_'+str(ch)+'_'
+    mySubsetOfFiles = []
+    for fileName in listOfAllFiles:
+        if containsPhrase in fileName:
+            mySubsetOfFiles.append(fileName)
+            listOfAllFiles.remove(fileName)
+
+    # read the cross correlations and average them
+    xCorr = np.load(mySubsetOfFiles[0])['arr_0']
+    xCorr = xCorr.astype(np.float32)
+    nt = xCorr.shape[1]
+    for i,fileName in enumerate(mySubsetOfFiles):
+        if i>0:
+            tempCorr = np.load(fileName)['arr_0']
+            xCorr = xCorr +  tempCorr.astype(np.float32)
+    xCorr = xCorr/len(mySubsetOfFiles)
+    
+    # symmetrize
+    xCorr = 0.5*(xCorr[:,nt/2:] + np.fliplr(xCorr[:,:nt/2+1]))
+    # normalize channel-wise
+    for i in range(startCh,endCh):
+        xCorr[i-startCh,:] = xCorr[i-startCh,:]/np.sum(np.absolute(xCorr[i-startCh,:]))
+
+    # *******NEEDS TO BE MADE MORE GENERAL******
+    secondsToPlot = 1.0
+    samplesPerSecond = 50
+    nSampsToPlot = int(samplesPerSecond*secondsToPlot)
+    xCorr = xCorr[:,:nSampsToPlot]
+    
+
+    # plot that average
+    clipVal = np.percentile(xCorr,99.8)
[…]

### 2017-04-27 code/src/plot_oneBitXCorrs_multi_file_lists.py

@@ -1,0 +1,114 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+paramsPath = sys.argv[1]
+sys.path.append(paramsPath)
+from params import *
+
+srcFile = open(srcChList,'r')
+srcChannelsStrings = srcFile.readlines()
+srcFile.close()
+srcChannels = [int(ch) for ch in srcChannelsStrings]
+
+# get the list of all the files
+listOfAllFiles = []
+for outfileListFile in outfileListFiles:
+    outFileList = open(outfileListFile,'r')
+    sublistOfAllFiles = outFileList.readlines()
+    sublistOfAllFiles = [(f.strip('\n'))+'.npz' for f in sublistOfAllFiles]
+    for s in sublistOfAllFiles:
+        listOfAllFiles.append(s)
+    outFileList.close()
+
+# read all the receiver locations
+recXs = np.zeros(endCh-startCh+1)
+recYs = np.zeros(endCh-startCh+1)
+recLocFile = open(recLocationsFilename,'r')
+firstChannel = True
+ch = startCh
+for line in recLocFile:
+    line = line.strip('\n')
+    stringList = line.split()
+    ch = int(stringList[0])
+    if (ch>=startCh) and (ch<=endCh):
+        recXs[ch-startCh] = float(stringList[1])
+        recYs[ch-startCh] = float(stringList[2])
+        # if it's the first channel to have the location read, have all earlier channels in same spot
+        if firstChannel:
+            for earlierChannel in range(startCh,ch):
+                recXs[earlierChannel-startCh] = recXs[ch-startCh]
+                recYs[earlierChannel-startCh] = recYs[ch-startCh]
+            firstChannel = False # no longer the first channel
+# after the last channel fill in rest to be same location
+while ch <= endCh:
+    recXs[ch-startCh] = recXs[ch-startCh-1]
+    recYs[ch-startCh] = recYs[ch-startCh-1]
+    ch = ch + 1
+
+for ch in srcChannels:
+    srcChX = recXs[ch-startCh]
+    srcChY = recYs[ch-startCh]
+
+    # get the list of file names for this source channel
+    containsPhrase = '_srcCh_'+str(ch)+'_'
+    mySubsetOfFiles = []
[…]

### 2017-04-27 code/src/reader.py

@@ -1,0 +1,9 @@
+#include <boost/python.hpp>
+
+BOOST_PYTHON_MODULE(reader)
+{
+	using namespace boost::python;
+	python::class_<READER>("READER")
+	.def(
+
+}

### 2017-04-27 code/src/params/params1/params.py
updated params for test data

@@ -1,13 +1,13 @@
 import datetime as dt
 
-startTime = dt.datetime(2016,9,4,5,0,54,932*1000)
+startTime = dt.datetime(2016,9,12,13,0,54,932*1000)
 secondsPerFile = 60
 secondsPerWindowWidth = 300
 secondsPerWindowOffset = 150
 xCorrMaxTimeLagSeconds = 3.0
-nFiles = 240
+nFiles = 9
 outfilePath = '/scratch/ermartin/oneBitXCorr/'
-outfileListFile = '/home/ermartin/StanfordDASArray/inter_results/outfileList1.txt'
+outfileListFile = '/home/CorrelationCode/src/inter_results/outfileList1.txt'
 srcChList = 'sourceList.txt'
 startCh = 15
 endCh = 300

### 2017-04-27 code/src/params/params1plot/params.py

@@ -1,0 +1,16 @@
+import datetime as dt
+
+startTime = dt.datetime(2016,9,4,5,0,54,932*1000)
+secondsPerFile = 60
+secondsPerWindowWidth = 300
+secondsPerWindowOffset = 150
+xCorrMaxTimeLagSeconds = 3.0
+nFiles = 240*6
+outfilePath = '/scratch/'
+outfileListFiles = ['/home/CorrelationCode/src/inter_results/outfileList1.txt']
+srcChList = 'sourceList.txt'
+startCh = 15
+endCh = 300
+minFrq = 0.2
+maxFrq = 24.0
+recLocationsFilename = '/home/CorrelationCode/src/inter_results/passive_channel_UTM_coords.txt'

### 2017-04-27 code/src/params/params1/params.py
fixed params1

@@ -6,7 +6,7 @@ secondsPerWindowWidth = 300
 secondsPerWindowOffset = 150
 xCorrMaxTimeLagSeconds = 3.0
 nFiles = 9
-outfilePath = '/scratch/ermartin/oneBitXCorr/'
+outfilePath = '/scratch/'
 outfileListFile = '/home/CorrelationCode/src/inter_results/outfileList1.txt'
 srcChList = 'sourceList.txt'
 startCh = 15

### 2017-04-27 code/src/params/params1plot/params.py
added array geometry and fixed plotting parameters

@@ -5,7 +5,7 @@ secondsPerFile = 60
 secondsPerWindowWidth = 300
 secondsPerWindowOffset = 150
 xCorrMaxTimeLagSeconds = 3.0
-nFiles = 240*6
+nFiles = 9
 outfilePath = '/scratch/'
 outfileListFiles = ['/home/CorrelationCode/src/inter_results/outfileList1.txt']
 srcChList = 'sourceList.txt'

### 2017-04-27 code/src/calc_oneBitXCorr.py
updated readme and packages in dockerfile

@@ -5,7 +5,7 @@ import datetime as dt
 import struct
 import obspy
 import sys
-import xcorr
+#import xcorr
 
 sys.path.append('/home/ermartin/PassiveSeismicArray')
 # get parameters: startTime, secondsPerFile, secondsPerWindowWidth, secondsPerWindowOffset, xCorrMaxTimeLagSeconds, nFiles, outfilePath, outfileList, srcChList, startCh, endCh, minFrq, maxFrq

### 2017-04-28 code/src/calc_oneBitXCorr.py
fixed some paths, and still working on Docker/Singularity but it runs on cees-tool-7 without containers

@@ -7,7 +7,7 @@ import obspy
 import sys
 #import xcorr
 
-sys.path.append('/home/ermartin/PassiveSeismicArray')
+#sys.path.append('/home/ermartin/PassiveSeismicArray')
 # get parameters: startTime, secondsPerFile, secondsPerWindowWidth, secondsPerWindowOffset, xCorrMaxTimeLagSeconds, nFiles, outfilePath, outfileList, srcChList, startCh, endCh, minFrq, maxFrq
 paramsPath = sys.argv[1]
 sys.path.append(paramsPath)

### 2017-04-28 code/src/params/params1/params.py

@@ -1,13 +1,13 @@
 import datetime as dt
 
-startTime = dt.datetime(2016,9,12,13,0,54,932*1000)
+startTime = dt.datetime(2016,9,12,12,50,54,932*1000)
 secondsPerFile = 60
 secondsPerWindowWidth = 300
 secondsPerWindowOffset = 150
 xCorrMaxTimeLagSeconds = 3.0
 nFiles = 9
 outfilePath = '/scratch/'
-outfileListFile = '/home/CorrelationCode/src/inter_results/outfileList1.txt'
+outfileListFile = 'inter_results/outfileList1.txt'
 srcChList = 'sourceList.txt'
 startCh = 15
 endCh = 300

### 2017-04-28 code/src/plot_oneBitXCorrs_multi_file_lists.py

@@ -5,6 +5,8 @@ paramsPath = sys.argv[1]
 sys.path.append(paramsPath)
 from params import *
 
+filenameStart = sys.argv[2]
+
 srcFile = open(srcChList,'r')
 srcChannelsStrings = srcFile.readlines()
 srcFile.close()
@@ -96,7 +98,7 @@ for ch in srcChannels:
     plt.colorbar()
     #plt.scatter(arrivalTimesV1[1:-1],channels[1:-1],s=0.5,c='k')
     #plt.scatter(arrivalTimesV2[1:-1],channels[1:-1],s=0.5,c='g')
-    plt.savefig('fig/xcorrRaster_noLines_1to18_vs_ch_'+str(ch)+'.pdf')
+    plt.savefig('fig/'+filenameStart+'_vs_ch_'+str(ch)+'.pdf')
     plt.clf()
 
     # plot some nearby wiggles

### 2017-04-28 code/src/plot_oneBitXCorrs_multi_file_lists.py
added some info about singularity but still trying to figure it out

@@ -94,6 +94,7 @@ for ch in srcChannels:
     plt.xlabel('time lag (sec)')
     plt.title('one bit cross-correlation, virtual source '+str(ch))
     plt.colorbar()
+    # uncomment two lines below if you want homogeneous velocity picks
     #plt.scatter(arrivalTimesV1[1:-1],channels[1:-1],s=0.5,c='k')
     #plt.scatter(arrivalTimesV2[1:-1],channels[1:-1],s=0.5,c='g')
     plt.savefig('fig/xcorrRaster_noLines_1to18_vs_ch_'+str(ch)+'.pdf')
